### 2019-01-25 this can be deleted.Rmd

---
title: "Untitled"
author: "DARS"
date: "1/24/2019"
output: pdf_document
---

```{r pressure, echo=FALSE}
test <- d_transcript %>%
  mutate(Round_grade = paste("g", round(Grade,0),sep=""),
         Values = T) %>%
  filter(!is.na(Round_grade), Grade>=0)

spread(test, key = "Round_grade", value = Values #, fill = F
               )%>%
  mutate(g1  = g0|g1,
         g2  = g1|g2,
         g3  = g2|g3,
         g4  = g3|g4,
         g5  = g4|g5,
         g6  = g5|g6,
         g7  = g6|g7,
         g8  = g7|g8,
         g9  = g8|g9,
         g10 = g9|g10,
         )%>%
  gather(key="Gr_string", value="Gr_bool", g0, g1, g2, g3, g4,g5,g6,g7,g8,g9,g10, na.rm=T) %>%
  arrange(`Course ID`, `Student ID`)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.