### 2019-02-18 Baby Recommender.Rmd

---
title: "Pillar A1"
author: "DARS"
date: "2/15/2019"
output: pdf_document
---

```{r library}
library(tidyverse)
```

#TO DO: prepare variables at begining

#CHECKING WHAT DATA SET TO USE
```{r}
(nrow(d_transcript)-nrow(d_transactions$taken_PF_HL))

s.1 <- data.frame(`Student ID` = unique(d_transactions$taken_PF_HL$sequenceID), first = 1)
s.2 <- data.frame(`Student ID` = unique(d_transcript$`Student ID`), first = 2)
(nrow(s.1)-nrow(s.2))

missing_students <- full_join(s.1,s.2, by = "Student.ID") %>% 
  filter(is.na(first.x)) %>%
  pull(Student.ID)

View(d_transcript%>%
  filter(`Student ID` %in% missing_students)) #students lost mostly only have external and core courses, there are a few internal courses missing: e.g. SSC1001

```

#Setup
This contains AR and SR
```{r setup}
load("App/rules.RDATA")
load("Output/transaction_df.RDATA")
```

First we gather a sample of students to provide recommendations to. We use `d_transactions$taken_PF_HL`
```{r sample}
students <- unique(d_transactions$taken_PF_HL$sequenceID)
size = floor(length(students)/10)

#
##Sample student IDs
set.seed(123)
s1 <- sample(
  x    = students,
  size = size 
)

#
##Transcripts of selected students
target_transcript <- d_transactions$taken_PF_HL %>%
  filter(
    sequenceID %in% s1
  ) %>%
  select(-`Period (additional)`, -`Period (additional bis)`, -`Concentration (additional)`) %>%
  select(sequenceID, eventID, item, item_PF, item_HL, Grade)

#c("Course ID", "Course Title.x", "Cluster", "n_ILO", "n_assessment", "Assessments Covered", "n_AoD", "AoD Covered", "Concentration", "Missing from ILO File", "Most Recent Catalogue", "Type", "Letters", "Number", "Level", "sequenceID", "Course Title.y", "Grade", "Year_numerical", "Period", "Number Repeated Attempt", "Academic Year", "Fail", "eventID", "PF", "HL", "item", "item_PF", "item_HL")

```

For all students
```{r truncated transcript}

course_sample_size <- 8

#
##Truncated prospective transcript
prospective_transcript <-target_transcript %>% 
  group_by(sequenceID) %>%
  top_n(course_sample_size, eventID)
#
##Truncated student transcript
sample_transcript <- prospective_transcript %>% 
  group_by(sequenceID) %>%
  top_n(5, eventID) %>%
  arrange(sequenceID)

#
##Only prospective courses
prospective_transcript <- prospective_transcript %>%
  anti_join(sample_transcript, by = c("sequenceID", "eventID", "item", "item_PF", "item_HL", "Grade"))  %>%
  arrange(sequenceID)

```

# TO DO: basic- doesn't look at confidence or lift, we need to do the filtering for meaningful rules first. Exclude taken already from the rhs.
##Matching rules to students:
###Taken
```{r will take}
# very basic, doesn't look at confidence or lift:
#(*) Do the filtering for meaningful rules first. 
#(**)Exclude taken already from the rhs.
taken_match <- AR$taken %>%
  #DO FILTERING FOR MEANINGFUL RULES HERE
  full_join(sample_transcript, by = c("lhs"="item"))%>%
  select(sequenceID, lhs, rhs) %>%
  drop_na() %>% #we lose 20 courses here -->20 rules or students are not matched with the transcripts
  group_by(sequenceID, rhs) %>%
  summarise(RHS_vote= n())  %>%
  arrange(sequenceID,desc(RHS_vote))

recommendations <- taken_match %>%
  group_by(sequenceID)         %>%
  filter(row_number() %in% 1:3) %>% #can use top_n
  arrange(sequenceID)          %>%
  select(sequenceID, RHS_vote, everything())

recommendations

```
####Check
```{r evaluation}
recommendations %>%
  full_join(prospective_transcript, 
            by= c("sequenceID"="sequenceID","rhs"="item")
            ) %>%
  drop_na()
```

###PF
```{r prospective red flag}
prospective_RedFlag <- AR$PF %>%
  left_join(prospective_courses, by = c("lhs"="item_PF"))%>%
  select(sequenceID, lhs, rhs) %>%
  drop_na() %>% #we eliminate rules that don't match students
  group_by(rhs) %>%
  mutate(RHS_vote= n())%>%
  arrange(desc(RHS_vote))

prospective_RedFlag
```
###HL
```{r expected low result}
expect_low_result <- AR$HL  %>%
  left_join(prospective_courses, by = c("lhs"="item_HL"))%>%
  select(sequenceID, lhs, rhs) %>%
  drop_na() %>% #we eliminate rules that don't match students
  group_by(rhs) %>%
  mutate(RHS_vote= n())%>%
  arrange(desc(RHS_vote))

expect_low_result
```

###
```{r merging all}
#
#setup: cleaning before merge
RedFlag <- prospective_RedFlag %>%
  select(sequenceID,lhs, rhs) %>%
  separate(rhs, into = c("prospective_item","expected_PF") ) %>%
  mutate(transcript_type = "RedFlag",
         Precedent_PF= lhs)%>%
  select(-lhs)

Low_Results <- expect_low_result %>%
  select(sequenceID,lhs, rhs) %>%
  separate(rhs, into = c("prospective_item","expected_HL") ) %>%
  mutate(transcript_type = "LowResults" ,
          Precedent_HL= lhs)%>%
  select(-lhs)

prosp_transcript <- prospective_transcript %>%
  select(sequenceID, item) %>%
  mutate(transcript_type = "StudentChoice" )

#
#Join
recommender <- recommendations %>%
  mutate(transcript_type ="Recommendations" #,
         #Reason = paste("votes = ", RHS_vote, sep = "") #this should be moved to the end and be calculated with a case_when
         ) %>% 
  full_join(prosp_transcript, by = c("sequenceID" = "sequenceID", "transcript_type" = "transcript_type", "rhs" = "item") )             %>%
  full_join(RedFlag,          by = c("sequenceID" = "sequenceID", "transcript_type" = "transcript_type", "rhs" = "prospective_item") ) %>%
  full_join(Low_Results,      by = c("sequenceID" = "sequenceID", "transcript_type" = "transcript_type", "rhs" = "prospective_item") ) %>%
  mutate(Reason = case_when(
   transcript_type=="Recommendations"~ RHS_vote,
   transcript_type=="RedFlag"        ~ expected_PF,
   transcript_type=="LowResults"     ~ expencted_,
   transcript_type=="StudentChoice"
  ))

```
##Matching sequences to students
###Taken
```{r will take}
d_suggestions<- SR$taken %>%
  left_join(sample_transcript, by = c("lhs"="item"))%>%
  drop_na() %>%
  group_by(rhs) %>%
  mutate(RHS_vote= n())%>%
  arrange(desc(RHS_vote))

recommendations <- taken_match[1:5,] %>% select(rhs)

```
####Check
```{r evaluation}
recommendations %>%
  left_join(prospective_courses, by= c("rhs"="Course ID")) %>%
  drop_na()
```

###PF
```{r prospective red flag}
d_prospective_RedFlag <- SR$PF %>%
  left_join(prospective_courses, by = c("lhs"="item_PF"))%>%
  drop_na() %>%
  group_by(rhs) %>%
  mutate(RHS_vote= n())%>%
  arrange(desc(RHS_vote))

d_prospective_RedFlag

```
###HL
```{r expected low result}
d_expect_low_result <- AR$HL  %>%
  left_join(prospective_courses, by = c("lhs"="item_HL"))%>%
  drop_na() %>%
  group_by(rhs) %>%
  mutate(RHS_vote= n())%>%
  arrange(desc(RHS_vote))

d_expect_low_result
```

TEST: For a single student:
```{r truncated transcript}

course_sample_size <- 5

student <- "6036972"

student_transcript <- target_transcript %>% 
    filter(sequenceID == student) %>%
  arrange(Year_numerical, Period)

length_student_transcript <- nrow(student_transcript)

#
##Truncated student transcript
sample_transcript      <- student_transcript[1:course_sample_size,]
prospective_courses <- student_transcript[(course_sample_size+1):(course_sample_size+2),] #%>% 
  #select(sequenceID, `Course ID`)
```