### 2019-02-19 test_delete_later_sofia.Rmd

---
title: "Topic Scores"
author: "DARS"
date: "2/18/2019"
output: pdf_document
---
#Set up
##Libraries
```{r libraries}
library(tidyverse)
```
##Raw Data
```{r setup, include=FALSE}
#
##d_course and d_transcript
load("Output/data_pillar_1.RDATA")

#
##LDA_models
load("Output/LDA_overview.RDATA")

#
##Topic Models test for ideal number of topics
load("Output/LDA_ntopics_description.RDATA")
load("Output/LDA_ntopics_overview.RDATA")

```

## Getting Functions (gamma and beta) 
```{r getting functions}
get_beta <- function(results){

  tidy(results, matrix = "beta") %>%
    mutate(topic = paste("Topic", topic)) %>%
      arrange(topic, desc(beta))

}

get_gamma <- function(results){

  tidy(results, matrix = "gamma") %>%
    mutate(topic = paste("Topic", topic)) %>%
      arrange(topic, desc(gamma))

}

```

##Getting applied (gamma and beta) 
```{r getting beta and gamma}
# k=10, 20, 35
beta_description <- lapply(
  LDA_model,
  get_beta
)

gamma_descriptions <- lapply(
  LDA_model,
  get_gamma
)

```
##Inspect topics
```{r inspect topics}
topics <- beta_description$k35 %>%
  group_by(topic) %>%
  top_n(10, beta)

#
##View
View(topics)

```
## simulating students and key words
```{r student simulation}
terms <- unique(beta_description$k35$term)

student <- list()
student$kw_policy <- c("international", "economic", "conflict", "develop", "policy", "war") #sample(terms, 5)
student$kw_climate <- c("climate", "sustainability", "change", "development", "develop","sustain", "sustainable") #sample(terms, 5)
```

##getting top courses per student
```{r topic score student}
standardize <- function (col) {
  col/sum(col)
}

#
##key word Contributions to Topic
 contribution_to_topics <- beta_description$k35                       %>%
  filter(term %in% student$kw_policy)       %>%
  arrange(topic, desc(beta))                %>%
  spread(key=topic, value=beta)             %>%
  mutate_at(.vars = vars(num_range("Topic ",1:35)),
            .funs = standardize)

```

##Topic contribution to course
```{r course_score}

#
## Topic contribution to course
gamma_descriptions$k35     %>%
  full_join(student$topic_score, by = "topic") %>%
  group_by(document, topic)                           %>%
  summarise(total_gamma= sum(gamma))
  mutate(X = gamma * B_proportion)             %>%
  group_by(document)                           %>%
  summarise(course_score = sum(X))             %>%
  arrange(desc(course_score))                  %>%
  left_join(
    select(d_course, `Course ID`, `Course Title`),
    by = c("document" = "Course ID")
    )

```
###View
```{r View course}
View(student$topic_score)
View(student$course_score)
```