### 2019-02-20 test_delete_later_sofia.Rmd
test key word selection

---
title: "Topic Scores"
author: "DARS"
date: "2/18/2019"
output: pdf_document
---
#Set up
##Libraries
```{r libraries}
library(tidyverse)
```

##Raw Data
```{r setup, include=FALSE}
#
##d_course and d_transcript
load("Output/data_pillar_1.RDATA")

#
##LDA_models
load("Output/LDA_overview.RDATA")

#
##Topic Models test for ideal number of topics
load("Output/LDA_ntopics_description.RDATA")
load("Output/LDA_ntopics_overview.RDATA")

```

## Getting Functions (gamma and beta) 
```{r getting functions}
get_beta <- function(results){

  tidy(results, matrix = "beta") %>%
    mutate(topic = paste("Topic", topic)) %>%
      arrange(topic, desc(beta))

}

get_gamma <- function(results){

  tidy(results, matrix = "gamma") %>%
    mutate(topic = paste("Topic", topic)) %>%
      arrange(topic, desc(gamma))

}

```

##Getting applied (gamma and beta) 
```{r getting beta and gamma}
# k=10, 20, 35
beta_description <- lapply(
  LDA_model,
  get_beta
)

gamma_descriptions <- lapply(
  LDA_model,
  get_gamma
)

```
##Inspect topics
```{r inspect topics}
topics <- beta_description$k35 %>%
  group_by(topic) %>%
  top_n(10, beta)

#
##View
View(topics)

```
## simulating students and key words
```{r student simulation}
terms <- unique(beta_description$k35$term)

student <- list()
student$kw_policy <- c("international", "economic", "conflict", "develop", "policy", "war", "philosophy") #sample(terms, 5)
student$kw_climate <- c("climate", "sustainability", "change", "development", "develop","sustain", "sustainable") #sample(terms, 5)

```

##getting top courses per student
```{r topic score student}
standardize <- function (col) {
  col/sum(col)
}

#
##key word Contributions to Topic
 contribution_to_topics #<- 
beta_description$k35                       %>%
  filter(term %in% student$kw_policy)                                 %>%
  arrange(topic, desc(beta))                                          %>%
  group_by(topic) %>%
  #this is new
  mutate(beta_sum=sum(beta),
         beta_rank=beta/beta_sum) %>%
  select(-beta_sum, -beta)
  #
  spread(key=topic, value=beta)                                       %>%
  mutate_at(.vars = vars(num_range("Topic ",1:35)),
            .funs = standardize)                                      %>%
   gather(key = "topic", value = "beta_rank", num_range("Topic ", 1:35))

  beta_description$k35                       %>%
  filter(term %in% student$kw_policy)                                 %>%
  arrange(topic, desc(beta))                                          %>%
  spread(key=topic, value=beta)                                       %>%
  mutate_at(.vars = vars(num_range("Topic ",1:35)),
            .funs = standardize)                                      %>%
   gather(key = "topic", value = "beta_rank", num_range("Topic ", 1:35))

```

##Topic contribution to course
```{r course_score}
#
##Extracts recommended courses
my_courses <- top_n(x = student$course_score, n = 5, wt =course_score) %>% 
  pull(document)

#
## Topic contribution to course
key_words <- gamma_descriptions$k35                             %>%
  filter(document %in% my_courses)                 %>%
  arrange(document, gamma)                         %>%
  full_join(contribution_to_topics, by = "topic" ) %>%
  arrange(document, desc(gamma), desc(beta_rank) ) %>%
  group_by(document #, 
           #gamma
           )                                       %>%
  filter(row_number() %in% 1:3)                    %>%  #makes most sense with just 1 kw # group_by(document) %>% filter(row_number() %in% 1:9) 
  group_by(document)                               %>%
  summarise(key_words = list(Reduce(f= c,
                               x = term)))

```
###View
```{r View course}
View(student$topic_score)
View(student$course_score)
```

###Selecting key words
```{r}
test <- beta_description$k35 %>%
  group_by(topic) %>%
  top_n(3, beta) %>%
  group_by(topic) %>%
  summarise(key_words= list(Reduce(f= c, x = term)))

kw <- unique(Reduce(x = test$key_words, f= c))

kw_stopwords =c("student", "skill", "academic", "bate", "participant", "gain", "primarily", "main", "2016", "action", "level", "issue", "concept", "close", "day", "edition", "primary", "project", "method", "apply", "busy", "address", "role", "post", "oxford", "cuss", "lecture", "pbl")

setdiff(kw, kw_stopwords)

```