### 2019-03-19 Pillar 2 - Topic Modeling.Rmd
get topic terms

---
title: "Pillar 2 - Topic Modeling"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Pillar 2/",
                      eval = FALSE)
```

```{r library, message = FALSE, eval = TRUE}
library(tidyverse)
library(tidytext)

library(ggwordcloud) # Word Clouds
library(topicmodels)
library(lemon)
library(ggthemes)
library(rlang)
library(ldatuning) # ideal number of topics in LDA

```

```{r function reorder_within & co., echo = FALSE}

# Function for ordering within facet:
# https://github.com/dgrtwo/drlib/blob/master/R/reorder_within.R

reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

scale_y_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_y_discrete(labels = function(x) gsub(reg, "", x), ...)
}
```

# Setup
We load the environment `output/data_pillar_2.RDATA`. It contains the data sets:
`d_AoD`, `d_assessment`, `d_course`, `d_description`, `d_manual`, `d_overview`, `d_transcript`, and a list `lists`.
For `d_description` and `d_overview`, we keep only the most recent year.
```{r import data, eval = TRUE}
load("output/data_pillar_2.RDATA")

# Only keep course descriptions and overviews from most recent year
d_text <- d_text %>% map(function(tb, y) tb %>% filter(year == max(year)))
```

# TF-IDF
## Functions for Generating Barplots and Word Clouds
The function `compute_tf_idf` computes the inverse term document frequency, we apply `compute_tf_idf` on descriptions, overviews and manuals. 
```{r compute_tf_idf, cache = TRUE}
#
##Function
compute_tf_idf <- function(data){

  data %>%
    count(
      `Course ID`,
      word
      ) %>%
    bind_tf_idf(
      term = word, 
      document = `Course ID`,
      n = n
      )

}

#
##Application
tf_idf <- map(d_text, .f = compute_tf_idf)

#
##Clean workspace
rm(compute_tf_idf)
```

The function `plot_tf_idf` plots a bar-graph of the most important words for each level of granularity and  their corresponding word clouds. 
```{r plot_tf_idf}
plot_tf_idf <- function(data, n_col = 5, id_plot = NULL){

  #
  # Barplot
  g <- data %>%
    ggplot(
      aes(
        x = reorder_within(word, tf_idf, facet), 
        y = tf_idf
        )
      ) +
    geom_col(
      show.legend = FALSE
      ) +
    labs(
      x = NULL,
      y = "tf-idf"
      ) +
    scale_x_reordered() +
    facet_wrap(
      facets = ~ facet, 
      scales = "free",
      ncol = n_col
      ) +
    coord_flip()

  ggsave(paste(id_plot, "_BP.jpeg", sep = ""), path = "Output/Plots/tf_idf", width = 16, height = 8)

  #
  # Word Cloud
  g <- data %>%
    group_by(
      facet
      ) %>%
    mutate(
      freq = tf_idf / sum(tf_idf)
      ) %>% # normalization within facet
    ggplot(
      aes(
        size = freq ^ 0.7, 
        label = word, 
        color = freq ^ 0.7
        )
      ) +
    geom_text_wordcloud_area(
      area_corr_power = 1,
      eccentricity    = 1,
      rm_outside      = T
      ) +
    scale_radius(
      range = c(2, 10),
      limits = c(0, NA)
      ) +
    scale_color_gradient(
      low = "red", 
      high = "blue"
      ) +
    facet_wrap(
      facet = ~ facet,
      ncol = n_col
      ) + 
    theme(
      strip.text.x     = element_text(),
      panel.background = element_rect(fill = "white"),
      plot.title       = element_text(hjust = 0.5)
    )

  ggsave(paste(id_plot, "_WC.jpeg", sep = ""), path = "Output/Plots/tf_idf", width = 16, height = 8)

}
```

## Course Level
The function `prepare_tf_idf_course` selects 25 random courses, the top ten words with highest tf_idf, and finally renames a facet to 'Course ID').
We apply the function to `id_plot` and pipe it to produce graphs.
```{r tf-idf course, cache = TRUE}
prepare_tf_idf_course <- function(data){

  data %>%

    # Selection 25 courses randomly
    filter(
      `Course ID` %in% sample(
        x = unique(.$`Course ID`),
        size = 25, 
        replace = FALSE
        )
      ) %>%

    # Select top 10 words per course 
    group_by(
      `Course ID`
      ) %>%
    filter(
      n >= 2
      ) %>%
    top_n(
      n = 10,
      wt = tf_idf
      ) %>%
    ungroup %>%

    # Prepare data for function `plot_tf_idf()`
    rename(
      facet = `Course ID`
      )

  }

set.seed(123)
tf_idf$overview %>%
  prepare_tf_idf_course %>%
  plot_tf_idf(id_plot = "Course_overview")

set.seed(123)
tf_idf$manual %>%
  prepare_tf_idf_course %>%
  plot_tf_idf(id_plot = "Course_manual")

rm(prepare_tf_idf_course)
```

## Cluster Level
```{r tf-idf cluster, cache = TRUE}
prepare_tf_idf_cluster <- function(data){

  data %>%

    # Include the variable Cluster
    left_join(
      select(d_course, `Course ID`, Cluster),
      by = "Course ID"
      ) %>%
    filter(
      !is.na(Cluster)
      ) %>%

    # Select top 10 words per cluster
    group_by(
      Cluster, 
      word
      ) %>%
    summarise(
      tf_idf = sum(tf_idf),
      n = sum(n)
      ) %>%
    filter(
      n >= 10
      ) %>%
    top_n(
      n = 10, 
      wt = tf_idf
      ) %>%
    ungroup %>%

    # Prepare data for function `plot_tf_idf()`
    rename(
      facet = Cluster
      )

}

tf_idf$overview %>%
  prepare_tf_idf_cluster %>%
  plot_tf_idf(id_plot = "Cluster_overview")

tf_idf$manual %>%
  prepare_tf_idf_cluster %>%
  plot_tf_idf(id_plot = "Cluster_manual")

rm(prepare_tf_idf_cluster)
```

## Concentration Level
```{r tf-idf concentration, cache = TRUE}
prepare_tf_idf_concentration <- function(data){

  data %>%

    # Include the variable Concentration
    left_join(
      select(d_course, `Course ID`, Concentration, `Concentration (additional)`),
      by = "Course ID"
      ) %>%
    filter(
      !is.na(Concentration)
      ) %>%
    gather(
      key = X,
      value = Concentration,
      Concentration, `Concentration (additional)`,
      na.rm = TRUE
      ) %>%

    # Select top 10 words per cluster
    group_by(
      Concentration, 
      word
      ) %>%
    summarise(
      tf_idf = sum(tf_idf),
      n = sum(n)
      ) %>%
    filter(
      n >= 10
      ) %>%
    top_n(
      n = 10,
      wt = tf_idf
      ) %>%
    ungroup %>%

    # Prepare data for function `plot_tf_idf()`
    rename(
      facet = Concentration
      )

}

tf_idf$overview %>%
  prepare_tf_idf_concentration %>%
  plot_tf_idf(id_plot = "Concentration_overview")

tf_idf$manual %>%
  prepare_tf_idf_concentration %>%
  plot_tf_idf(id_plot = "Concentration_manual")

rm(prepare_tf_idf_concentration, plot_tf_idf)
```

# LDA
##Setup
###Cast DTM
The function `my_cast_dtm` casts our dataframes into a DTM.
```{r my_cast_dtm, eval = TRUE}
my_cast_dtm <- function(data) data %>%
  count(`Course ID`, word) %>%
  cast_dtm(`Course ID`, word, n)

d_cast <- d_text %>% map(my_cast_dtm)

rm(my_cast_dtm)
```
###Controls
The control parameters we use are:
```{r controls, eval = TRUE}
my_control <- list(

  nstart = 10,
  seed   = 1 : 10,
  best   = TRUE,

  burnin = 1000,
  iter   = 15000,
  thin   = 100

)
```
The control parameters we use for fast testing are:
```{r fast control}
my_control_fast <- list(

  nstart = 1,
  seed   = 1,
  best   = TRUE,

  burnin = 5,
  iter   = 10,
  thin   = 2

)
```

##Ideal Number of Topics Package Test
To decide the number of topics to discover, we run three different tests.
The function `my_FindTopicsNumber` builds a Topic Model and applies the measurements 
```{r function my_FindTopicsNumber }
my_FindTopicsNumber <- function(data_cast, n_topics){

  FindTopicsNumber(

  dtm = data_cast,
  topics = n_topics,
  metrics = c("CaoJuan2009", "Arun2010", "Deveaud2014"),
  method = "Gibbs",
  control = my_control,
  mc.cores = 4L

  )

}
```
We run the test for descriptions, documenting the duration (starting_time - ending_time). We build 95 models (5 to 100 in increments of 1).
```{r description topics, cache = TRUE}
start.time <- Sys.time()

result_description <- d_description_cast %>% my_FindTopicsNumber(n_topics = seq(from = 5, to = 100, by = 2))

end.time <- Sys.time()
time.taken <- end.time - start.time

save(result_description,
     file = "Output/LDA_ntopics.RDATA")

print(time.taken)
```

We run the test for overview, documeting its duration. 
```{r overview topics, cache = TRUE}
start.time <- Sys.time()

result_overview <- d_overview_cast %>% my_FindTopicsNumber(n_topics = seq(from = 5, to = 100, by = 2))

end.time <- Sys.time()
time.taken <- end.time - start.time

save(result_overview , result_description,
     file = "Output/LDA_ntopics.RDATA")

print(time.taken)
```

We run the test for manuals, documenting its duration.
```{r manual topics, cache = TRUE}
start.time <- Sys.time()

result_manual <- d_manual_cast %>% my_FindTopicsNumber(n_topics = seq(from = 5, to = 75, by = 5))

end.time <- Sys.time()
time.taken <- end.time - start.time

save(result_manual, result_overview , result_description,
     file = "Output/LDA_ntopics.RDATA")

print(time.taken)
```

from 20, to 40, by 5 topics takes
22min for overview
11min for description
5.5hrs for manuals

We examine the result of the experiment by plotting them on a graph.
```{r LDA number topics plot}

load("Output/LDA_ntopics.RDATA")

result_description %>%
  FindTopicsNumber_plot

result_overview %>%
  FindTopicsNumber_plot

result_manual %>%
  FindTopicsNumber_plot

```
From the results we see that the ideal number of topics is somewhere between 35 and 40.

## Fitting Model
The function `my_LDA` fits an LDA model on the DTM using the previous controls.
```{r my_LDA, eval= TRUE}
my_LDA <- function(n_topics, corpus, control = my_control){

  LDA_model <- LDA(
    x       = corpus,
    k       = n_topics,
    method  = "Gibbs",
    control = control
    )
}

```

We train the LDA for different numbers of topics (5 to 250 by 5), we save the resulting 50 models under `topic_model`
###Functions for models (get gamma, get beta, extract_dtm)
```{r getting gamma and beta, eval = TRUE}
#extract dtm
extract_dtm <- function(origin) d_cast[[origin]]

#extract Distributions (Gamma/Beta)
get_distributions <- function(model, distrib_str){ #input gamma or beta
  tidy(model, matrix = distrib_str)        %>%
    mutate(topic = paste("Topic", topic) ) %>%
      arrange(topic, desc(!!sym(distrib_str)))
}

#perplexity Function
my_perplexity <- function(model, dtm){
  perplexity(model, dtm, estimate_theta = FALSE)
}

```
###Training Models
```{r LDA fit, cache = TRUE, eval = TRUE}

test_topic_model <- tibble( #change to topic_model
  #repeat each topic number twice
  n_topic = rep(seq(from = 2, to = 4, by = 1), each = 2) # from = 5, to = 250, by = 5

  ) %>%
  mutate(Origin     = rep(c("overview", "manuals"), 3),         # should be 50
         `Topic ID` = paste(Origin, n_topic, sep = "_"),
         DTM        = pmap(.l = list(Origin), .f = extract_dtm),

         #fit model
         Model  = pmap(.l = list(n_topic, DTM), .f = my_LDA, control = my_control_fast),

         #Gamma distribution
         Gamma = pmap(.l = list(Model), .f = get_distributions, distrib_str = "gamma"),

         #Beta distribution
         Beta = pmap(.l = list(Model),  .f = get_distributions, distrib_str = "beta"),

         #Perplexity
         Perplexity = pmap_dbl(.l = list(Model, DTM), .f = my_perplexity),

         #Loglikelihood
         LogLikelihood = pmap_dbl(.l = list(Model), .f = logLik) #this removes the degrees of freedom data
         )

#View(test_topic_model)
test_topic_model
```
#DELETE LATERRRRRR- HELPER IMPORT
```{r}
load("Output/topic_model.RDATA")
```

##Model Evaluation
###Plots
```{r perplexity plots, eval = TRUE}
#Function
plot_evaluation <- function(origin){

  #data
  test_topic_model %>%

    #Wrangling
    select(n_topic, Origin, Perplexity, LogLikelihood) %>%
    gather(key = "Eval Type", value = "Eval Value", c(Perplexity, LogLikelihood)) %>%
    filter(Origin == origin) %>%

    #Graphics
    ggplot(mapping = aes(x = n_topic, y = `Eval Value`)) +
    geom_point() +
    geom_line() +
    facet_grid(rows = vars(`Eval Type`), scales = "free") +
    theme_minimal()

}

#Plot
plot_evaluation("overview")
plot_evaluation("manuals")
```

##Saving model_data:
For internal use:
```{r}
# save(topic_model, file = "Output/topic_model.RDATA" ) #needs to be renamed to match what is currently test_topic_model
```

#Distributions
#NOTE: select how many topics to use

##Distributions of best models
```{r Distributions, eval = FALSE}
#select number of topics
n_topics_overview <- 36
n_topics_manual   <- 40

extract_overview <- function() topic_model_gb %>% filter(n_topic == n_topics_overview)
extract_manual   <- function() topic_model_gb %>% filter(n_topic == n_topics_manual  )

df %>% mutate(var_new = list(var_old1, var_old2) %>% pmap(my_function))

data <- tribble(
  ~ ID,
  "overview",
  "manual"
  ) %>%
  mutate(df = if_else(ID == "overview", extract_overview(), extract_manual))

#Crop models
TM_overview <- topic_model_gb %>%
  filter(n_topic== n_topics_overview)

TM_manual   <- topic_model_gb %>%
  filter(n_topic== n_topics_manual)

#Distributions
distribution <- list()

distribution$beta$overview  <- TM_overview %>% pull(b_overview)
distribution$beta$manual    <- TM_manual   %>% pull(b_manual)

distribution$gamma$overview <- TM_overview %>% pull(g_overview)
distribution$gamma$manual   <- TM_manual   %>% pull(g_manual)

```
##Inspect topics
```{r inspect topics, eval = FALSE}

# FUNCTION
#get_top_match <- function(<INPUT GAMMA OR BETA DISTRIB>){<FUNCTION GENERALIZATION>}
get_top_match <- function(d_distribution){
  print(!!sym(d_distribution))
}
Hi <- "hello"
get_top_match(Hi)

get_top_terms_topic <- function(d_beta){
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(10, beta) %>%
    select(topic, term) %>%
    group_by(topic) %>%
    summarise(top_words = list(term))

}

get_top_topics_course <- function(d_gamma){
  data.frame(d_gamma) %>%
    group_by(document) %>%
    top_n(5, gamma) %>%
    select(document, topic) %>%
    summarise(top_topic = list(topic))
}

get_kw <- function(d_beta){
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(10, beta) %>%
    select(topic, term) %>%
    group_by(topic) %>%
    summarise(top_words = list(term)) %>%
    select(top_words) %>%
    collapse(top_words)
}
#Saving to models
test_topic_model <-  test_topic_model %>%
  mutate(top_terms_topic = map(Beta, .f = get_top_terms_topic))

test_topic_model <- test_topic_model %>%
  mutate(top_topic_course = map(Gamma, .f = get_top_topics_course))

```
#App
##Selectig keywords
```{r key words}
get_top_words <- function(d_beta, n = 5){
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(n, beta) %>%
    pull(term)%>%
    unique
}

go_down <- function(my_list, i){
  my_list[[i]]
}

get_key_words <-  function(d_top_words){

  select(top_words) %>%
    top_n(5)
}

View(test_topic_model)

a <- test_topic_model[["top_terms_topic"]]

d_top_words <- a[1]

#mutate(kw = get_key_words(top_terms_topic)) 
d_top_words <- a$top_terms_topic

kw <- lapply(
  distribution$beta,
  get_top_words
)

```
####TEMPORTARY: stopwords for keywords
This section is for presentations
```{r stopwords for kw}
#kw$overview
not_include <- c("source", "primary", "type", "assess", "involve", "e.g", "discussion", "institute", "ha", "introduce",
                 "role","address", "student", "examination","coordinator", "maastrichtuniversity.nl", NA, "identify", "manner", "mechanism", "concept", "hour", "teach",  "depth","question", "specific", "busy","academic", "semester", "week", "test", "session", "team", "communicate", "theory", "text", "library", "article", "examine", "current", "routledge", "temporary", "flow", "level", "study")

remove_words <- function(x){
  setdiff(x, not_include)
}

kw1 <- lapply(
  kw,
  remove_words
)

```
##Convenience vectors for APP
```{r for convenience}
# set seed
set.seed(1)

#
## All courses
course_all <- d_course %>%
  inner_join(
    d_transcript,
    by = "Course ID"
  ) %>%
  distinct(
    `Course ID`
  ) %>%
  # remove semester abroad, skills and projects
  filter(
    ! str_detect(`Course ID`, pattern = "SA|SKI|PRO")
  ) %>%
  pull

#
##courses following_semester
course_following_semester <- sample(
  x       = course_all,
  size    = 60,
  replace = FALSE
  ) %>% sort
```
## Saving files for APP
```{r app files}
save(distribution, kw, course_all, course_following_semester,
     file = "App/Recommender System/data_topic_models.RDATA")
```

# Visualization
### Functions 
This part is adapted 
```{r LDA prepare, eval = F}
prepare_data_LDA_beta <- function(d_beta){

  data.frame(d_beta) %>%
    mutate(topic = paste("Topic", topic)) %>%
      group_by(topic) %>%
        top_n(10, beta) %>%
      ungroup %>%
      arrange(topic, desc(beta))

} #repetition??? of inspect topics?

prepare_data_LDA_gamma <- function(d_gamma, level = Course){

  data_gamma <- data.frame(d_gamma) %>%
    mutate(topic = paste("Topic", topic)) %>%
    rename(`Course ID` = document)

  if(ensym(level) == sym("Course")) data_gamma <- data_gamma %>%
                                      rename(facet = `Course ID`)

  if(ensym(level) == sym("Cluster")) data_gamma <- data_gamma %>% 
                                       left_join(select(d_course, `Course ID`, Cluster), by = "Course ID") %>%
                                       filter(!is.na(Cluster)) %>%
                                       rename(facet = Cluster)

  if(ensym(level) == sym("Concentration")) data_gamma <- data_gamma %>%
                                             left_join(select(d_course, `Course ID`, Concentration, `Concentration (additional)`), by = "Course ID") %>%
                                             gather(X, Concentration, Concentration, `Concentration (additional)`, na.rm = TRUE) %>%
                                             rename(facet = Concentration)

  data_gamma <- data_gamma %>%
    group_by(facet, topic) %>%
      summarise(gamma = sum(gamma)) %>%
    filter(gamma > 0.05) %>%
    ungroup %>%
    arrange(facet, desc(gamma))

  return(data_gamma)

}

# Bet Distribution
test <- list()

a <-  names(distribution)[1]
test[[a]] <- lapply(distribution[[a]],
                      prepare_data_LDA_beta)

# Gamma Distribution
b <- names(distribution)[2]
test[[b]] <- lapply(distribution[[b]],
                    prepare_data_LDA_gamma)

rm(a,b #,test
   )
```
This part is not adapted
```{r LDA visualize, eval = F}
visualize_LDA_beta <- function(data_prepared, id_plot = "test"){

  data_prepared %>%
  ggplot(aes(x = reorder_within(term,
                                by = beta,
                                within = topic),
             y = beta,
             fill = topic)) +
    geom_col(show.legend = F) +
    facet_wrap(~ topic, scales = "free") +
    scale_x_reordered() +
    labs(x = "Terms", y = "Beta", title = "Main Terms of each Topic") +
    coord_flip() +
    theme_light() +
    theme(plot.title = element_text(hjust = 0.5))

  ggsave(paste(id_plot, "beta.jpeg"),
         width = 16, height = 10, path = "Plots/Topic Model")

}

visualize_LDA_gamma1 <- function(data_prepared, id_plot = "test"){

  data_prepared %>%
    ggplot(aes(reorder_within(facet, 
                              by = gamma, 
                              within = topic), 
               y = gamma, 
               fill = topic)) +
    geom_col(show.legend = F) +
    facet_wrap(~ topic, 
               scales = "free") +
    scale_x_reordered() +
    coord_flip() +
   # labs(x = ensym(facet), y = "Gamma", title = paste("Main", ensym(facet), "of each Topic") +
    theme_light() +
    theme(axis.text.x = element_text(angle = 90, hjust = 0),
          plot.title = element_text(hjust = 0.5))

  ggsave(paste(id_plot, "gamma1.jpeg"),
         width = 16, height = 10, path = "Plots/Topic Model")

}

visualize_LDA_gamma2 <- function(data_prepared, id_plot = "test"){

  data_prepared %>%
    ggplot(aes(reorder_within(topic, 
                              by = gamma,
                              within = facet), 
               y = gamma, 
               fill = topic)) +
    geom_col(show.legend = F) +
    facet_wrap(~ facet, 
               scales = "free") +
    scale_x_reordered() +
    coord_flip() +
  #  labs(x = "Topics", y = "Gamma", title = "Main Topics of each Courses/Cluster") +
    theme_light() +
    theme(axis.text.x = element_text(angle = 90, hjust = 0),
          plot.title = element_text(hjust = 0.5))

  ggsave(paste(id_plot, "gamma2.jpeg"),
         width = 16, height = 10, path = "Plots/Topic Model")

}
```

```{r LDA visualize all distrib, eval = F}
visualize_LDA_all_distrib <- function(data, level = Course, id_plot = "test"){

  # Beta distribution
  data %>% 
    prepare_data_LDA_beta %>%
    visualize_LDA_beta(id_plot = id_plot)

  # Gamma 1 distribution
  data %>% 
    prepare_data_LDA_gamma(level = !!ensym(level)) %>%
    visualize_LDA_gamma1(id_plot = id_plot)

  # Gamma 2 distribution
  data %>% 
    prepare_data_LDA_gamma(level = !!ensym(level)) %>%
    visualize_LDA_gamma2(id_plot = id_plot)

}
```

```{r LDA visualize all level, eval = F}
visualize_LDA_all_level <- function(data, id_plot = "test"){

  data %>% 
    visualize_LDA_all_distrib(level = Course,
                              id_plot = paste(id_plot, "_course"))

  data %>% 
    visualize_LDA_all_distrib(level = Cluster,
                              id_plot = paste(id_plot, "_cluster"))

  data %>% 
    visualize_LDA_all_distrib(level = Concentration,
                              id_plot = paste(id_plot, "_concentration"))

}
```

### Plots
```{r LDA plot, cache = TRUE, eval = F}
load("LDA.RDATA")

LDA_description_5 %>% visualize_LDA_all_level(id_plot = "description_k5")

LDA_overview_5 %>% visualize_LDA_all_level(id_plot = "overview_k5")

LDA_manual_5 %>% visualize_LDA_all_level(id_plot = "manual_k5")
```

```{r label, eval = FALSE, eval = F}
tidy(LDA_manual_10, "beta") %>%
  filter(beta > 1e-3) %>%
  bind_tf_idf(term = term, document = topic, n = beta) %>%
  group_by(topic) %>%
  top_n(10, tf_idf) %>%
  arrange(topic, desc(tf_idf))

  count(term, topic) %>%
  arrange(desc(n)) %>%

  filter(beta > 1e-3) %>%
  group_by(term) %>%
    mutate(n_topic = n(),
           beta = beta / n_topic) %>%
  group_by(topic) %>%
    top_n(10, beta)

  count(topic, term, wt = beta) %>%
  (term = term, document = topic, n = n) %>%
  group_by(topic) %>%
  top_n(10, tf_idf)

  group_by(term) %>%
  summarise(beta = sum(beta)) %>%
  arrange(desc(beta))

labels <- list(
  description_10 = "X"
)
```

```{r label bis, eval = FALSE}
visualize_LDA(LDA_cluster_10, id_plot = "Cluster, k = 10 with labelled topics", facet = T,
              topic_names = c("Psychology & Information", "Economics and Development", "Biology", "Chemistry and Math",
                              "Sociology", "Law", "Research", "International Politics",
                              "Philosophy", "Humanities"))

visualize_LDA(LDA_cluster_17, id_plot = "Cluster, k = 17 with labelled topics", facet = T,
              topic_names = c("Science/Knowledge", "Psycho", "Biology", "Chemistry", "Culture",
                              "Conflict", "Research", "Foreign Policy", "Arts", "Literature",
                              "Communication Skills", "Law", "Humanities", "Computer", "Economics", 
                              "Math", "Management and ?"))
visualize_LDA(LDA_cluster_25, id_plot = "Cluster, k = 25 with labelled topics", facet = T,
              topic_names = c(
                "Information", "Psycho", "Bio", "Chem", "Sociology",
                "History", "skills", "Europe", "Philosophy", "IR",
                "Conference", "Law", "Media", "Philosophy", "Economics",
                "Math", "Management and Literature", "Entrepreneurship", "Conflict", "Research",
                "Arts", "Computer Science", "Policy?", "Physiology", "Qual. Res."))
```