### 2019-03-21 Pillar 2 - Topic Modeling.Rmd

---
title: "Pillar 2 - Topic Modeling"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Pillar 2/",
                      eval = FALSE)
```

#Setup
##Libraries
```{r library, message = FALSE, eval = TRUE}
library(tidyverse)
library(tidytext)

library(ggwordcloud) # Word Clouds
library(topicmodels)
library(lemon)
library(ggthemes)
library(rlang) #quote
library(ldatuning) # ideal number of topics in LDA

```

##Functions for internal faceting
```{r function reorder_within & co., echo = FALSE, eval = TRUE}

# Function for ordering within facet:
# https://github.com/dgrtwo/drlib/blob/master/R/reorder_within.R

reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

scale_y_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_y_discrete(labels = function(x) gsub(reg, "", x), ...)
}
```

##Loading data
We load the environment `output/data_pillar_2.RDATA`. It contains the data sets:
`d_course`, `d_transcript`, and a list `d_text` with the textual data for catalogues and manuals (for $overview, $description, $manuals we have (`Course ID`, `year`, `word_original`, `word`) ).

For `d_text` we keep only the most recent year.
```{r import data, eval = TRUE}
#load
load("output/data_pillar_2.RDATA")

# keep most recent year
d_text <- d_text %>% map(function(tb, y) tb %>% filter(year == max(year)))
```

# TF-IDF
We will perform an initial (visual) exploratory analysis of our textual data using TF- IDF.

## Functions for Generating Barplots and Word Clouds
The function `compute_tf_idf` computes the inverse term document frequency, we apply `compute_tf_idf` on descriptions, overviews and manuals. 
```{r compute_tf_idf, cache = TRUE}
#
##Function
compute_tf_idf <- function(data){

  data %>%
    count(
      `Course ID`,
      word
      ) %>%
    bind_tf_idf(
      term = word, 
      document = `Course ID`,
      n = n
      )
}

#
##Application
tf_idf <- map(d_text, .f = compute_tf_idf)

#
##Clean workspace
rm(compute_tf_idf)
```

The function `plot_tf_idf` plots a bar-graph of the most important words for each level of granularity and  their corresponding word clouds. 
```{r plot_tf_idf}
plot_tf_idf <- function(data, n_col = 5, id_plot = NULL){

  #
  # Barplot
  g <- data %>%
    ggplot(
      aes(
        x = reorder_within(word, tf_idf, facet), 
        y = tf_idf
        )
      ) +
    geom_col(
      show.legend = FALSE
      ) +
    labs(
      x = NULL,
      y = "tf-idf"
      ) +
    scale_x_reordered() +
    facet_wrap(
      facets = ~ facet, 
      scales = "free",
      ncol = n_col
      ) +
    coord_flip()

  ggsave(paste(id_plot, "_BP.jpeg", sep = ""), path = "Output/Plots/tf_idf", width = 16, height = 8)

  #
  # Word Cloud
  g <- data %>%
    group_by(
      facet
      ) %>%
    mutate(
      freq = tf_idf / sum(tf_idf)
      ) %>% # normalization within facet
    ggplot(
      aes(
        size = freq ^ 0.7, 
        label = word, 
        color = freq ^ 0.7
        )
      ) +
    geom_text_wordcloud_area(
      area_corr_power = 1,
      eccentricity    = 1,
      rm_outside      = T
      ) +
    scale_radius(
      range = c(2, 10),
      limits = c(0, NA)
      ) +
    scale_color_gradient(
      low = "red", 
      high = "blue"
      ) +
    facet_wrap(
      facet = ~ facet,
      ncol = n_col
      ) + 
    theme(
      strip.text.x     = element_text(),
      panel.background = element_rect(fill = "white"),
      plot.title       = element_text(hjust = 0.5)
    )

  ggsave(paste(id_plot, "_WC.jpeg", sep = ""), path = "Output/Plots/tf_idf", width = 16, height = 8)

}
```

## Course Level
The function `prepare_tf_idf_course` selects 25 random courses, the top ten words with highest tf_idf, and finally renames a facet to 'Course ID').
We apply the function to `id_plot` and pipe it to produce graphs.
```{r tf-idf course, cache = TRUE}
prepare_tf_idf_course <- function(data){

  data %>%

    # Selection 25 courses randomly
    filter(
      `Course ID` %in% sample(
        x = unique(.$`Course ID`),
        size = 25, 
        replace = FALSE
        )
      ) %>%

    # Select top 10 words per course 
    group_by(
      `Course ID`
      ) %>%
    filter(
      n >= 2
      ) %>%
    top_n(
      n = 10,
      wt = tf_idf
      ) %>%
    ungroup %>%

    # Prepare data for function `plot_tf_idf()`
    rename(
      facet = `Course ID`
      )

  }

set.seed(123)
tf_idf$overview %>%
  prepare_tf_idf_course %>%
  plot_tf_idf(id_plot = "Course_overview")

set.seed(123)
tf_idf$manual %>%
  prepare_tf_idf_course %>%
  plot_tf_idf(id_plot = "Course_manual")

rm(prepare_tf_idf_course)
```

## Cluster Level
```{r tf-idf cluster, cache = TRUE}
prepare_tf_idf_cluster <- function(data){

  data %>%

    # Include the variable Cluster
    left_join(
      select(d_course, `Course ID`, Cluster),
      by = "Course ID"
      ) %>%
    filter(
      !is.na(Cluster)
      ) %>%

    # Select top 10 words per cluster
    group_by(
      Cluster, 
      word
      ) %>%
    summarise(
      tf_idf = sum(tf_idf),
      n = sum(n)
      ) %>%
    filter(
      n >= 10
      ) %>%
    top_n(
      n = 10, 
      wt = tf_idf
      ) %>%
    ungroup %>%

    # Prepare data for function `plot_tf_idf()`
    rename(
      facet = Cluster
      )

}

tf_idf$overview %>%
  prepare_tf_idf_cluster %>%
  plot_tf_idf(id_plot = "Cluster_overview")

tf_idf$manual %>%
  prepare_tf_idf_cluster %>%
  plot_tf_idf(id_plot = "Cluster_manual")

rm(prepare_tf_idf_cluster)
```

## Concentration Level
```{r tf-idf concentration, cache = TRUE}
prepare_tf_idf_concentration <- function(data){

  data %>%

    # Include the variable Concentration
    left_join(
      select(d_course, `Course ID`, Concentration, `Concentration (additional)`),
      by = "Course ID"
      ) %>%
    filter(
      !is.na(Concentration)
      ) %>%
    gather(
      key = X,
      value = Concentration,
      Concentration, `Concentration (additional)`,
      na.rm = TRUE
      ) %>%

    # Select top 10 words per cluster
    group_by(
      Concentration, 
      word
      ) %>%
    summarise(
      tf_idf = sum(tf_idf),
      n = sum(n)
      ) %>%
    filter(
      n >= 10
      ) %>%
    top_n(
      n = 10,
      wt = tf_idf
      ) %>%
    ungroup %>%

    # Prepare data for function `plot_tf_idf()`
    rename(
      facet = Concentration
      )

}

tf_idf$overview %>%
  prepare_tf_idf_concentration %>%
  plot_tf_idf(id_plot = "Concentration_overview")

tf_idf$manual %>%
  prepare_tf_idf_concentration %>%
  plot_tf_idf(id_plot = "Concentration_manual")

rm(prepare_tf_idf_concentration, plot_tf_idf)
```

# LDA
##Setup
###Cast DTM
The function `my_cast_dtm` casts our dataframes into a DTM.
```{r my_cast_dtm, eval = TRUE}
my_cast_dtm <- function(data) data %>%
  count(`Course ID`, word) %>%
  cast_dtm(`Course ID`, word, n)

d_cast <- d_text %>% map(my_cast_dtm)

rm(my_cast_dtm)
```
###Controls
The control parameters we use are:
```{r controls, eval = TRUE}
my_control <- list(

  nstart = 10,
  seed   = 1 : 10,
  best   = TRUE,

  burnin = 1000,
  iter   = 15000,
  thin   = 100

)
```
The control parameters we use for fast testing are:
```{r fast control}
my_control_fast <- list(

  nstart = 1,
  seed   = 1,
  best   = TRUE,

  burnin = 5,
  iter   = 10,
  thin   = 2

)
```

##Ideal Number of Topics Package Test
To decide the number of topics to discover, we run three different tests.
###Function
The function `my_FindTopicsNumber` builds a Topic Model and applies the measurements 
```{r function my_FindTopicsNumber }
my_FindTopicsNumber <- function(data_cast, n_topics){

  FindTopicsNumber(

  dtm = data_cast,
  topics = n_topics,
  metrics = c("CaoJuan2009", "Arun2010", "Deveaud2014"),
  method = "Gibbs",
  control = my_control,
  mc.cores = 4L

  )

}
```
###Train
```{r package test on topics, cache = TRUE}
#Start timer
start.time <- Sys.time()

#
##apply my_FindTopicsNumber
result_topic_test <- as_tibble(map(.x = d_cast, .f = my_FindTopicsNumber, n_topics = seq(from = 5, to = 100, by = 2)))

#Stop timer
end.time <- Sys.time()
print(end.time - start.time)

#
##Save results
#save(result_topic_test,
#     file = "Output/LDA_ntopics.RDATA") #this file still contains old results.
```
Previous separate experiments (from 20, to 40, by 5 topics) had the following time:
* Overview:     22min 
* Description:  11min 
* Manuals:      5.5hrs

###Graph 
We examine the result of the experiment by plotting them on a graph.
```{r LDA number topics plot}
#load("Output/LDA_ntopics.RDATA")

result_topic_test %>% map(.f = ldatuning::FindTopicsNumber_plot) #order is: Overview, Manuals, Description
```
From the results we see that the ideal number of topics is somewhere between 35 and 40.

## Fitting Model
The function `my_LDA` fits an LDA model on the DTM using the previous controls.
```{r my_LDA, eval= TRUE}
my_LDA <- function(n_topics, corpus, control = my_control){

  LDA_model <- LDA(
    x       = corpus,
    k       = n_topics,
    method  = "Gibbs",
    control = control
    )
}

```

###Functions for models (extract_dtm, get_distributions, perplexity)
The following functions are used as input in a map to compute and evaluate topic models. 
```{r getting gamma and beta, eval = TRUE}
#extract dtm
extract_dtm <- function(origin) d_cast[[origin]]

#extract Distributions (Gamma/Beta)
get_distributions <- function(model, distrib_str){ #input gamma or beta
  tidy(model, matrix = distrib_str)        %>%
    mutate(topic = paste("Topic", topic) ) %>%
      arrange(topic, desc(!!sym(distrib_str)))
}

#perplexity Function
my_perplexity <- function(model, dtm){
  perplexity(model, dtm, estimate_theta = FALSE)
}

```

The following functions allow us to inspect the results from our models(inspect distributions and get kw):
```{r}
 # get_top_match <- function(d_distribution){
 # 
 #  #save user input as string
 #  name_selected <- deparse(substitute(d_distribution)) #takes Gamma or Beta
 #  beta_gamma_sym <- tolower(name_selected)             #transforms to "gamma" or "beta"
 # 
 #  #Compute Case (Beta or Gamma, respectively)
 #  grouping         <- ifelse( name_selected == "Beta",  "topic", "document")
 #  objects_to_group <- ifelse( name_selected == "Beta",  "term",  "topic"   )
 #  top_num          <- ifelse( name_selected == "Beta",   10,      5       ) #select how many terms or topics(respectively) to show
 # 
 #  #
 #  ## DATA
 #  data.frame(d_distribution) %>%
 #    group_by(!!sym(grouping)) %>%
 #    top_n(top_num, !!sym(beta_gamma_sym)) %>%
 #    select(!!sym(grouping), !!sym(objects_to_group)) %>%
 #    group_by(!!sym(grouping)) %>%
 #    summarise(top = list(!!sym(objects_to_group)))
 # }

get_top_terms_topic <- function(d_beta){
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(10, beta) %>%
    select(topic, term) %>%
    group_by(topic) %>%
    summarise(top_words = list(term))

}

get_top_topics_course <- function(d_gamma){
  data.frame(d_gamma) %>%
    group_by(document) %>%
    top_n(5, gamma) %>%
    select(document, topic) %>%
    summarise(top_topic = list(topic))
}

get_kw <- function(d_beta){
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(3, beta) %>%
    ungroup()%>%
    pull(term)%>%
    unique()
}

```
###Training Models
We train the LDA for different numbers of topics (5 to 250 by 5), we save the resulting 50 models under `topic_model`, along with the DTM used to build the model, the corpus of origin (overview or manuals), the word and topic distributions results from that model, it's evaluation measures, and vectors of words and topics that allow us humans to get a feel of the model, along with some key words.
```{r LDA fit, cache = TRUE, eval = TRUE}

test_topic_model <- tibble( #CHANGE NAME TO: topic_model

  #repeat each topic number twice (once for overview and once for manual)
  n_topic = rep(seq(from = 2, to = 4, by = 1), each = 2) # from = 5, to = 250, by = 5

  ) %>%
  mutate(Origin     = rep(c("overview", "manuals"), 3),         # should be 50
         `Topic ID` = paste(Origin, n_topic, sep = "_"),

         #Document Term Matrix
         DTM        = pmap(.l = list(Origin), .f = extract_dtm),

         #Fit Model
         Model  = pmap(.l = list(n_topic, DTM), .f = my_LDA, control = my_control_fast),

         #Gamma distribution
         Gamma = pmap(.l = list(Model), .f = get_distributions, distrib_str = "gamma"),

         #Beta distribution
         Beta = pmap(.l = list(Model),  .f = get_distributions, distrib_str = "beta"),

         #Perplexity
         Perplexity = pmap_dbl(.l = list(Model, DTM), .f = my_perplexity),

         #Loglikelihood
         LogLikelihood = pmap_dbl(.l = list(Model), .f = logLik), #this removes the degrees of freedom data

         #
         ##Inspect

         #top words per topic
         top_terms_topic = map(Beta, .f = get_top_terms_topic),

         #top topics per course
         top_topic_course = map(Gamma, .f = get_top_topics_course),

         #Select keywords
         kw = map(Beta, .f = get_kw)
         )

# Clean workspace

# rm(extract_dtm, get_distributions, my_perplexity,
#    get_top_terms_topic, get_top_topics_course, get_kw)

# test_topic_model
```

##Plots - Model Evaluation
To evaluate how many topics are optiomal, we graph the perplexity and log Likelihood of models against the number of topics.
```{r perplexity plots, eval = TRUE}
#Function
plot_evaluation <- function(origin){

  #data
  test_topic_model %>%

    #Wrangling
    select(n_topic, Origin, Perplexity, LogLikelihood) %>%
    gather(key = "Eval Type", value = "Eval Value", c(Perplexity, LogLikelihood)) %>%
    filter(Origin == origin) %>%

    #Graphics
    ggplot(mapping = aes(x = n_topic, y = `Eval Value`)) +
    geom_point() +
    geom_line() +
    facet_grid(rows = vars(`Eval Type`), scales = "free") +
    theme_minimal()

}

#Plot
plot_evaluation("overview")
plot_evaluation("manuals")
```

##Saving model_data:
We save our tibble with all the information.
```{r}
# save(topic_model, file = "Output/topic_model.RDATA" ) #needs to be renamed to match what is currently test_topic_model
```

#App
For our recommender system, we pass the distributions, number of topics, origin, and key words to the app.
##Distributions and Key words
The function `clean_kw` removes stopwords from our key words.
```{r function clean key words}
clean_kw <- function(kw_list){

  #stopwords
  kw_stopwords <- c("source", "primary", "type", "assess", "involve", "e.g", "discussion", "institute", "ha", "introduce",
                 "role","address", "student", "examination","coordinator", "maastrichtuniversity.nl", "identify", "manner", "mechanism", "concept", "hour",
                 "teach",  "depth","question", "specific", "busy","academic", "semester", "week", "test", "session", "team", "communicate", "theory", "text",
                 "library", "article", "examine", "current", "routledge", "temporary", "flow", "level", "study", "task")

  #remove stopwords
  kw <- setdiff(kw_list, kw_stopwords)
  #return
  return(kw[!is.na(kw)])

}

```
On our `topic_model` tibble we filter only information relevant for the app, and apply `clean_kw` on our `kw` column
```{r crop and clean TM for App}
app_topic_model <- test_topic_model %>% select(`Topic ID`, n_topic, Origin, Gamma, Beta, kw) %>%
  mutate(kw = map(kw, .f = clean_kw))
```

##Convenience Vectors
For the app we also need to have a list of all courses and an example of a prospective semester
```{r for convenience}
# set seed
set.seed(1)

#
## All courses
course_all <- d_course %>%
  inner_join(
    d_transcript,
    by = "Course ID"
  ) %>%
  distinct(
    `Course ID`
  ) %>%
  # remove semester abroad, skills and projects
  filter(
    ! str_detect(`Course ID`, pattern = "SA|SKI|PRO")
  ) %>%
  pull

#
##courses following_semester
course_following_semester <- sample(
  x       = course_all,
  size    = 60,
  replace = FALSE
  ) %>% sort

```
## Saving files for APP
```{r app files}
# save(app_topic_model, course_all, course_following_semester,
#      file = "App/Recommender System/data_topic_models.RDATA")
```
# Visualization
### Functions 
This part is adapted 
```{r LDA prepare, eval = F}
prepare_data_LDA_beta <- function(d_beta){

  data.frame(d_beta) %>%
    mutate(topic = paste("Topic", topic)) %>%
      group_by(topic) %>%
        top_n(10, beta) %>%
      ungroup %>%
      arrange(topic, desc(beta))

} #repetition??? of inspect topics?

prepare_data_LDA_gamma <- function(d_gamma, level = Course){

  data_gamma <- data.frame(d_gamma) %>%
    mutate(topic = paste("Topic", topic)) %>%
    rename(`Course ID` = document)

  if(ensym(level) == sym("Course")) data_gamma <- data_gamma %>%
                                      rename(facet = `Course ID`)

  if(ensym(level) == sym("Cluster")) data_gamma <- data_gamma %>% 
                                       left_join(select(d_course, `Course ID`, Cluster), by = "Course ID") %>%
                                       filter(!is.na(Cluster)) %>%
                                       rename(facet = Cluster)

  if(ensym(level) == sym("Concentration")) data_gamma <- data_gamma %>%
                                             left_join(select(d_course, `Course ID`, Concentration, `Concentration (additional)`), by = "Course ID") %>%
                                             gather(X, Concentration, Concentration, `Concentration (additional)`, na.rm = TRUE) %>%
                                             rename(facet = Concentration)

  data_gamma <- data_gamma %>%
    group_by(facet, topic) %>%
      summarise(gamma = sum(gamma)) %>%
    filter(gamma > 0.05) %>%
    ungroup %>%
    arrange(facet, desc(gamma))

  return(data_gamma)

}

# Bet Distribution
test <- list()

a <-  names(distribution)[1]
test[[a]] <- lapply(distribution[[a]],
                      prepare_data_LDA_beta)

# Gamma Distribution
b <- names(distribution)[2]
test[[b]] <- lapply(distribution[[b]],
                    prepare_data_LDA_gamma)

rm(a,b #,test)
```
This part is not adapted
```{r LDA visualize, eval = F}
visualize_LDA_beta <- function(data_prepared, id_plot = "test"){

  data_prepared %>%
  ggplot(aes(x = reorder_within(term,
                                by = beta,
                                within = topic),
             y = beta,
             fill = topic)) +
    geom_col(show.legend = F) +
    facet_wrap(~ topic, scales = "free") +
    scale_x_reordered() +
    labs(x = "Terms", y = "Beta", title = "Main Terms of each Topic") +
    coord_flip() +
    theme_light() +
    theme(plot.title = element_text(hjust = 0.5))

  ggsave(paste(id_plot, "beta.jpeg"),
         width = 16, height = 10, path = "Plots/Topic Model")

}

visualize_LDA_gamma1 <- function(data_prepared, id_plot = "test"){

  data_prepared %>%
    ggplot(aes(reorder_within(facet, 
                              by = gamma, 
                              within = topic), 
               y = gamma, 
               fill = topic)) +
    geom_col(show.legend = F) +
    facet_wrap(~ topic, 
               scales = "free") +
    scale_x_reordered() +
    coord_flip() +
   # labs(x = ensym(facet), y = "Gamma", title = paste("Main", ensym(facet), "of each Topic") +
    theme_light() +
    theme(axis.text.x = element_text(angle = 90, hjust = 0),
          plot.title = element_text(hjust = 0.5))

  ggsave(paste(id_plot, "gamma1.jpeg"),
         width = 16, height = 10, path = "Plots/Topic Model")

}

visualize_LDA_gamma2 <- function(data_prepared, id_plot = "test"){

  data_prepared %>%
    ggplot(aes(reorder_within(topic, 
                              by = gamma,
                              within = facet), 
               y = gamma, 
               fill = topic)) +
    geom_col(show.legend = F) +
    facet_wrap(~ facet, 
               scales = "free") +
    scale_x_reordered() +
    coord_flip() +
  #  labs(x = "Topics", y = "Gamma", title = "Main Topics of each Courses/Cluster") +
    theme_light() +
    theme(axis.text.x = element_text(angle = 90, hjust = 0),
          plot.title = element_text(hjust = 0.5))

  ggsave(paste(id_plot, "gamma2.jpeg"),
         width = 16, height = 10, path = "Plots/Topic Model")

}
```

```{r LDA visualize all distrib, eval = F}
visualize_LDA_all_distrib <- function(data, level = Course, id_plot = "test"){

  # Beta distribution
  data %>% 
    prepare_data_LDA_beta %>%
    visualize_LDA_beta(id_plot = id_plot)

  # Gamma 1 distribution
  data %>% 
    prepare_data_LDA_gamma(level = !!ensym(level)) %>%
    visualize_LDA_gamma1(id_plot = id_plot)

  # Gamma 2 distribution
  data %>% 
    prepare_data_LDA_gamma(level = !!ensym(level)) %>%
    visualize_LDA_gamma2(id_plot = id_plot)

}
```

```{r LDA visualize all level, eval = F}
visualize_LDA_all_level <- function(data, id_plot = "test"){

  data %>% 
    visualize_LDA_all_distrib(level = Course,
                              id_plot = paste(id_plot, "_course"))

  data %>% 
    visualize_LDA_all_distrib(level = Cluster,
                              id_plot = paste(id_plot, "_cluster"))

  data %>% 
    visualize_LDA_all_distrib(level = Concentration,
                              id_plot = paste(id_plot, "_concentration"))

}
```

### Plots
```{r LDA plot, cache = TRUE, eval = F}
load("LDA.RDATA")

LDA_description_5 %>% visualize_LDA_all_level(id_plot = "description_k5")

LDA_overview_5 %>% visualize_LDA_all_level(id_plot = "overview_k5")

LDA_manual_5 %>% visualize_LDA_all_level(id_plot = "manual_k5")
```

```{r label, eval = FALSE, eval = F}
tidy(LDA_manual_10, "beta") %>%
  filter(beta > 1e-3) %>%
  bind_tf_idf(term = term, document = topic, n = beta) %>%
  group_by(topic) %>%
  top_n(10, tf_idf) %>%
  arrange(topic, desc(tf_idf))

  count(term, topic) %>%
  arrange(desc(n)) %>%

  filter(beta > 1e-3) %>%
  group_by(term) %>%
    mutate(n_topic = n(),
           beta = beta / n_topic) %>%
  group_by(topic) %>%
    top_n(10, beta)

  count(topic, term, wt = beta) %>%
  (term = term, document = topic, n = n) %>%
  group_by(topic) %>%
  top_n(10, tf_idf)

  group_by(term) %>%
  summarise(beta = sum(beta)) %>%
  arrange(desc(beta))

labels <- list(
  description_10 = "X"
)
```

```{r label bis, eval = FALSE}
visualize_LDA(LDA_cluster_10, id_plot = "Cluster, k = 10 with labelled topics", facet = T,
              topic_names = c("Psychology & Information", "Economics and Development", "Biology", "Chemistry and Math",
                              "Sociology", "Law", "Research", "International Politics",
                              "Philosophy", "Humanities"))

visualize_LDA(LDA_cluster_17, id_plot = "Cluster, k = 17 with labelled topics", facet = T,
              topic_names = c("Science/Knowledge", "Psycho", "Biology", "Chemistry", "Culture",
                              "Conflict", "Research", "Foreign Policy", "Arts", "Literature",
                              "Communication Skills", "Law", "Humanities", "Computer", "Economics", 
                              "Math", "Management and ?"))
visualize_LDA(LDA_cluster_25, id_plot = "Cluster, k = 25 with labelled topics", facet = T,
              topic_names = c(
                "Information", "Psycho", "Bio", "Chem", "Sociology",
                "History", "skills", "Europe", "Philosophy", "IR",
                "Conference", "Law", "Media", "Philosophy", "Economics",
                "Math", "Management and Literature", "Entrepreneurship", "Conflict", "Research",
                "Arts", "Computer Science", "Policy?", "Physiology", "Qual. Res."))
```