### 2019-03-28 Pillar 2 -(STM) Topic Modeling.Rmd
Template for (STM) Topic Model, runs until evaluation

---
title: "Pillar 2 - (STM) Topic Modeling"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: TRUE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Pillar 2/",
                      eval = FALSE)
```

#Setup
##Libraries
```{r library, message = FALSE, eval = TRUE}
library(tidyverse)
library(tidytext)

#Topic Models
library(stm)
library(topicmodels)
library(quanteda) #used for DTM implementation called dfm

library(ggwordcloud) # Word Clouds

library(lemon)
library(ggthemes)
library(rlang) #quote
library(ldatuning) # ideal number of topics in LDA

```

##Functions for internal faceting
```{r function reorder_within & co., echo = FALSE, eval = TRUE}

# Function for ordering within facet:
# https://github.com/dgrtwo/drlib/blob/master/R/reorder_within.R

reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

scale_y_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_y_discrete(labels = function(x) gsub(reg, "", x), ...)
}
```

##Loading data
We load the environment `output/data_pillar_2.RDATA`. It contains the data sets:
`d_course`, `d_transcript`, and a list `d_text` with the textual data for catalogues and manuals (for $overview, $description, $manuals we have (`Course ID`, `year`, `word_original`, `word`) ).

For `d_text` we keep only the most recent year.
```{r import data, eval = TRUE}
#load
load("output/data_pillar_2.RDATA")

# keep most recent year
d_text <- d_text %>% map(function(tb, y) tb %>% filter(year == max(year)))
```

# TF-IDF
We will perform an initial (visual) exploratory analysis of our textual data using TF- IDF.

## Functions for Generating Barplots and Word Clouds
The function `compute_tf_idf` computes the inverse term document frequency, we apply `compute_tf_idf` on descriptions, overviews and manuals. 
```{r compute_tf_idf, cache = TRUE}
#
##Function
compute_tf_idf <- function(data){

  data %>%
    count(
      `Course ID`,
      word
      ) %>%
    bind_tf_idf(
      term = word, 
      document = `Course ID`,
      n = n
      )
}

#
##Application
tf_idf <- map(d_text, .f = compute_tf_idf)

#
##Clean workspace
rm(compute_tf_idf)
```

The function `plot_tf_idf` plots a bar-graph of the most important words for each level of granularity and  their corresponding word clouds. 
```{r plot_tf_idf}
plot_tf_idf <- function(data, n_col = 5, id_plot = NULL){

  #
  # Barplot
  g <- data %>%
    ggplot(
      aes(
        x = reorder_within(word, tf_idf, facet), 
        y = tf_idf
        )
      ) +
    geom_col(
      show.legend = FALSE
      ) +
    labs(
      x = NULL,
      y = "tf-idf"
      ) +
    scale_x_reordered() +
    facet_wrap(
      facets = ~ facet, 
      scales = "free",
      ncol = n_col
      ) +
    coord_flip()

  ggsave(paste(id_plot, "_BP.jpeg", sep = ""), path = "Output/Plots/tf_idf", width = 16, height = 8)

  #
  # Word Cloud
  g <- data %>%
    group_by(
      facet
      ) %>%
    mutate(
      freq = tf_idf / sum(tf_idf)
      ) %>% # normalization within facet
    ggplot(
      aes(
        size = freq ^ 0.7, 
        label = word, 
        color = freq ^ 0.7
        )
      ) +
    geom_text_wordcloud_area(
      area_corr_power = 1,
      eccentricity    = 1,
      rm_outside      = T
      ) +
    scale_radius(
      range = c(2, 10),
      limits = c(0, NA)
      ) +
    scale_color_gradient(
      low = "red", 
      high = "blue"
      ) +
    facet_wrap(
      facet = ~ facet,
      ncol = n_col
      ) + 
    theme(
      strip.text.x     = element_text(),
      panel.background = element_rect(fill = "white"),
      plot.title       = element_text(hjust = 0.5)
    )

  ggsave(paste(id_plot, "_WC.jpeg", sep = ""), path = "Output/Plots/tf_idf", width = 16, height = 8)

}
```

## Course Level
The function `prepare_tf_idf_course` selects 25 random courses, the top ten words with highest tf_idf, and finally renames a facet to 'Course ID').
We apply the function to `id_plot` and pipe it to produce graphs.
```{r tf-idf course, cache = TRUE}
prepare_tf_idf_course <- function(data){

  data %>%

    # Selection 25 courses randomly
    filter(
      `Course ID` %in% sample(
        x = unique(.$`Course ID`),
        size = 25, 
        replace = FALSE
        )
      ) %>%

    # Select top 10 words per course 
    group_by(
      `Course ID`
      ) %>%
    filter(
      n >= 2
      ) %>%
    top_n(
      n = 10,
      wt = tf_idf
      ) %>%
    ungroup %>%

    # Prepare data for function `plot_tf_idf()`
    rename(
      facet = `Course ID`
      )

  }

set.seed(123)
tf_idf$overview %>%
  prepare_tf_idf_course %>%
  plot_tf_idf(id_plot = "Course_overview")

set.seed(123)
tf_idf$manual %>%
  prepare_tf_idf_course %>%
  plot_tf_idf(id_plot = "Course_manual")

rm(prepare_tf_idf_course)
```

## Cluster Level
```{r tf-idf cluster, cache = TRUE}
prepare_tf_idf_cluster <- function(data){

  data %>%

    # Include the variable Cluster
    left_join(
      select(d_course, `Course ID`, Cluster),
      by = "Course ID"
      ) %>%
    filter(
      !is.na(Cluster)
      ) %>%

    # Select top 10 words per cluster
    group_by(
      Cluster, 
      word
      ) %>%
    summarise(
      tf_idf = sum(tf_idf),
      n = sum(n)
      ) %>%
    filter(
      n >= 10
      ) %>%
    top_n(
      n = 10, 
      wt = tf_idf
      ) %>%
    ungroup %>%

    # Prepare data for function `plot_tf_idf()`
    rename(
      facet = Cluster
      )

}

tf_idf$overview %>%
  prepare_tf_idf_cluster %>%
  plot_tf_idf(id_plot = "Cluster_overview")

tf_idf$manual %>%
  prepare_tf_idf_cluster %>%
  plot_tf_idf(id_plot = "Cluster_manual")

rm(prepare_tf_idf_cluster)
```

## Concentration Level
```{r tf-idf concentration, cache = TRUE}
prepare_tf_idf_concentration <- function(data){

  data %>%

    # Include the variable Concentration
    left_join(
      select(d_course, `Course ID`, Concentration, `Concentration (additional)`),
      by = "Course ID"
      ) %>%
    filter(
      !is.na(Concentration)
      ) %>%
    gather(
      key = X,
      value = Concentration,
      Concentration, `Concentration (additional)`,
      na.rm = TRUE
      ) %>%

    # Select top 10 words per cluster
    group_by(
      Concentration, 
      word
      ) %>%
    summarise(
      tf_idf = sum(tf_idf),
      n = sum(n)
      ) %>%
    filter(
      n >= 10
      ) %>%
    top_n(
      n = 10,
      wt = tf_idf
      ) %>%
    ungroup %>%

    # Prepare data for function `plot_tf_idf()`
    rename(
      facet = Concentration
      )

}

tf_idf$overview %>%
  prepare_tf_idf_concentration %>%
  plot_tf_idf(id_plot = "Concentration_overview")

tf_idf$manual %>%
  prepare_tf_idf_concentration %>%
  plot_tf_idf(id_plot = "Concentration_manual")

rm(prepare_tf_idf_concentration, plot_tf_idf)
```

# STM
##Setup
###Cast DTM
The function `my_cast_dtm` casts our dataframes into a DTM (dfm from package quanteda)
```{r my_cast_dtm, eval = TRUE}
my_cast_dtm <- function(data) data %>%
  count(`Course ID`, word) %>%
  cast_dfm(`Course ID`, word, n)

d_cast <- d_text %>% map(my_cast_dtm)

rm(my_cast_dtm)
```

###STM Corpus
We convert our DTMs into STM corpus to access to full functionality in STM package.
```{r as STM corpus, eval = TRUE}
stm_corpus <- d_cast %>% map(asSTMCorpus)
```

##Fitting STM
We will to train several models to have some flexibility in the number of topics we select. For this purpose we want to create two tibbles. One that trains on the entire data (for use), and one that we divide on training set and test set. 
###Functions
Several functions are needed to work with tibble and tidyverse functionality:
####My_stm Function
The function `my_LDA` fits an LDA model on the DTM using the previous controls.
```{r my_STM, eval= TRUE}
my_stm <- function(n_topics, corpus){

  STM_model <- stm(corpus$documents, corpus$vocab, K = n_topics, init.type = "Spectral")
}
```
####My_stm_eval Function
This function applied stm but with heldout vocab and documents. 
```{r}
my_stm_eval <- function(heldout, n_topic){
  stm::stm(heldout$documents, heldout$vocab, n_topic, init.type = "Spectral")
}
```

####Extraction Functions
We will be working with many lists of lists. The following functions help us extract information from lower levels of the list (e.g. access element in second list).
```{r getting gamma and beta, eval = TRUE}
#
##From external Lists:

#Extract DTM
extract_dtm <- function(origin) d_cast[[origin]]
#Extract CORPUS
extract_corpus <- function(origin) stm_corpus[[origin]]

#
## From STM_corpus class

#Extract documents
extract_documents <- function(corpus) corpus$documents
#Extract vocab
extract_vocab     <- function(corpus) corpus$vocab
#Extract data
extract_data     <- function(corpus) corpus$data

#
##From Partition

#Extract heldout docs
extract_held_doc <- function(heldout) heldout$documents#[[1]]
#Extract herlout vocab
extract_held_vocab <- function(heldout) heldout$vocab#[[1]]
#Extract missing
extract_helout_missing <- function(heldout) heldout$missing#[[1]]

#
## MODEL HELDOUT
my_model <- function(heldout){

  docs <- heldout$documents[[1]]
  voc  <- heldout$vocab
  k    <- n_topics

  Model <- stm(docs, voc, 5, init.type = "Spectral")

  return(Model)
}

#
##From Model
#tidy gamma
my_tidy_gamma <- function(Gamma, DTM){
  tidy(Gamma, matrix = "gamma", document_names = rownames(DTM))}

#extract

#extract Distributions (Gamma/Beta)
get_distributions <- function(model, distrib_str){ #input gamma or beta
  tidy(model, matrix = distrib_str)        %>%
    mutate(topic = paste("Topic", topic) ) %>%
      arrange(topic, desc(!!sym(distrib_str)))}

```
####Inspection functions
The following functions allow us to inspect the results from our models(inspect distributions and get kw):
```{r}
get_top_terms_topic <- function(d_beta){
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(10, beta) %>%
    select(topic, term) %>%
    group_by(topic) %>%
    summarise(top_words = list(term))

}

get_top_topics_course <- function(d_gamma){
  data.frame(d_gamma) %>%
    group_by(document) %>%
    top_n(5, gamma) %>%
    select(document, topic) %>%
    summarise(top_topic = list(topic))
}

get_kw <- function(d_beta){
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(3, beta) %>%
    ungroup()%>%
    pull(term)%>%
    unique()
}
```
###Model Tibbles
####For use:
```{r models in tibble}
topic_numbers <- seq(from = 5, to = 10, by = 5) # seq(from = 5, to = 20, by = 5) #seq(from = 5, to = 150, by = 5)

stm_model <- tibble(
  #repeat each topic number twice (once for overview and once for manual)
  n_topic = rep(topic_numbers, each = 2) ) %>%

  mutate(
    #Identification information
    Origin     = rep(c("overview", "manuals"), length(topic_numbers)),         
    `Topic ID` = paste(Origin, n_topic, sep = "_"),

    #DTM
    DTM       = pmap(.l = list(Origin), .f = extract_dtm),
    doc_names = map(DTM, rownames),

    #STM Corpus
    Corpus           = pmap(.l = list(Origin), .f = extract_corpus), 
    #Documents- STM corpus
    Documents_stm     = pmap(.l=list(Corpus),    .f = extract_documents),
    #Vocab- STM corpus
    Vocab_stm         = pmap(.l=list(Corpus),    .f = extract_vocab),
    #Data- STM corpus
    Data_stm          = pmap(.l=list(Corpus),    .f = extract_data),

    #Fit Model 
    Model  = pmap(.l = list(n_topic, Corpus), .f = my_stm),

    #Evaluation
    Exclusivity        = map(Model, exclusivity),
    semantic_coherence = pmap(.l = list(Model,DTM), semanticCoherence, M = 10), #M = num. of top words per topic
    #Alretnative input to semantic_coherence = pmap(.l = list(Model, Documents_stm), semanticCoherence, M = 10),

    #
    ##Inspection

    #Beta distribution
    Beta = pmap(.l = list(Model),  .f = tidy),
    #Gamma distribution
    Gamma = pmap(.l = list(Model, DTM), .f = my_tidy_gamma), #fix this

    #top words per topic
    top_terms_topic = map(Beta, .f = get_top_terms_topic),

    #top topics per course
    top_topic_course = map(Gamma, .f = get_top_topics_course),

    #Select keywords
    kw = map(Beta, .f = get_kw)

    #END MUTATE PARENTHESIS
         )

```

```{r} 
#testing 
test_model <- stm_eval_model[1, ]

my_stm_eval(list(test_model$documents, test_model$vocab, test_model$n_topic))

my_stm_eval(documents = test_model$documents, vocab = test_model$vocab, n_topic = test_model$n_topic)

test <- list(test_model$documents, test_model$vocab, test_model$n_topic)
test %>% my_stm_eval

test <- list(stm_eval_model$documents, stm_eval_model$vocab, stm_eval_model$n_topic)

test %>% my_stm_eval

test <- stm_eval_model %>% mutate(test = list(documents, vocab, n_topic) %>% pmap(list))

stm_eval_model %>% mutate(test = test %>% map(my_stm_eval))
test1<- test$test[[1]]

```

####For evaluation
```{r models in tibble}
topic_numbers <-  seq(from = 5, to = 10, by = 5) #seq(from = 5, to = 150, by = 5)

#
##STM EVALUATION TIBBLE
stm_eval_model <- tibble(

  #repeat each topic number twice (once for overview and once for manual)
  n_topic = rep(topic_numbers, each = 2)

  ) %>%
  mutate(
    #Identification information
    Origin     = rep(c("overview", "manuals"), length(topic_numbers)),         
    `Topic ID` = paste(Origin, n_topic, sep = "_"),

    #DTM
    DTM       = pmap(.l = list(Origin),  .f = extract_dtm),
    Heldout   = pmap(.l = list(DTM),     .f = make.heldout, proportion = 0.5),

    #Fitting Model
    Model     = pmap(.l = list(Heldout, n_topic), .f = my_stm_eval),

    #Evaluation measures
    Exclusivity        = map(Model, exclusivity),

    semantic_coherence = pmap(.l = list(Model, Heldout), 
                              function(model,  heldout) semanticCoherence(model, heldout$documents)),

    eval_heldout       = pmap(.l = list(Model, Heldout), 
                              function(model,  heldout) eval.heldout(model, heldout$missing)),

    residual           = pmap(.l = list(Model, Heldout), 
                              function(model,  heldout) checkResiduals(model, heldout$documents, tol = 1/100)),

    bound              = pmap_dbl(.l = list(Model), 
                                  function(model) max(model$convergence$bound)),

    lfact              = map_dbl(Model,       
                                 function(model) lfactorial(model$settings$dim$K)),

    lbound             = bound + lfact,
    iterations         = map_dbl(Model,       
                                 function(model) length(model$convergence$bound))
    ) 
```

##Finding Number of Topics
We want to estimate the number of topics that exist in the data, to select a model for them.

###Using search K
`stm::searchK` helps in the search.
The function `findK` is used to vectorise searchK so we can apply it to the different elements of our corpus list (overview, manuals, description) in one go.
```{r findK}
#findK function 
findK <- function(dtm, k_topics){
  searchK(dtm$documents, dtm$vocab, K = k_topics, init.type = "Spectral")
}
```

We train models for 74 different numbers of topics (14-200 by 2).
Time estimate: for seq(from = 50, to = 100, by = 10) #18 models trained took 4:31-3:09 = (1 hour and 22 min)
```{r finding K}
#select number of topics
k_topics = seq(from = 4, to = 150, by = 2)

#Run findK
findingK <- stm_corpus %>%
  map(findK, k_topics)
```
We inspect the findings:
```{r plot evaluation}
findK_plots <- map(findingK, plot) #see how to show.
```

```{r save findK}
#save files
save(findingK, file= "STM_estimate_k.RDATA")

#clean
rm(findK, k_topics)
```
###Manual selection