### 2019-03-28 Data Preparation.Rmd
Prepare text data in a separate RMD file for clarity

---
title: "Data Preparation"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Data Preparation/")

library(tidyverse)
library(tidytext)
library(gsheet) # import spreadsheets from google drive
library(tm)
library(hunspell) # Stemmer
library(magrittr)
```

# Import Data
The datasets we use are saved as spreadsheet on our google drive *DARS* (with exeption of grade data saved as csv files on the computer for privacy reasons). We use the function `gsheet2tbl` to import them to `R Studio` as tibbles. We use the tibble data format (an evolution of the data frame format) because this is the format of reference of the `tidyverse` on whose tools our analysis is heavily based.

## List of AoD's and Assessments
First, we import the spreadsheet with information pertraining the Aims of the Degree (AoD) and Assessments from the drive and save it under `lists_brut`. ` lists_brut` contains 4 columns, under which we find the `19` types of assessment, the `18` aims of the degree (AoD) of the degree, and  two columns containing binary vectors indicating which assessment types and AoD  we will consider when ploting the data[^1].
Then we create a list with this same columns, but instead of having binary vectors for the plots, we keep vectors of only the names of relevant assessments and AoDs for the plots (`Assessment_plot`adnd `AoD_plot` respectively). We also any imported emtpy cells.
```{r list_AoD_assessment}
list_AoD_assessment <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1239912347') %>%

  map(na.omit)

# selection of most important types of assessment an AoD to keep plots clear
list_AoD_assessment$`Assessment for Plot` <- list_AoD_assessment$Assessment[list_AoD_assessment$`Assessment for Plot` == 1] %T>% print
list_AoD_assessment$`Aim for Plot`        <- list_AoD_assessment$Aim       [list_AoD_assessment$`Aim for Plot`        == 1] %T>% print
```

## Course Data
We import three spreadsheets from the drive and transform them into the so-called *tidy format*. The tibble `d_course` contains information at the course-level such as in which cluster and concentration(s) they belong, and in which period(s) they are offered. The tibble `d_assessment` indicates which type(s) of assessment each course contains with one row per course-assessment; and the tibble `d_ILO` indicates which AoD(s) the intended learning objectives (ILOs) of the courses cover with one row per course-ILO-AoD.

```{r d_course}
d_course <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1655700848') %>% print
```

```{r d_assessment}
d_assessment <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1102665750') %>%

  mutate_if(is.numeric, as.logical) %>%

  # Transform data into tidy format to facilitate manipulation
  gather(key   = Assessment, value = assessment_covered, Paper : Participation) %>%

  filter(assessment_covered) %>%

  select(- c(assessment_covered, `Comment on Assessment`)) %T>%

  print
```

```{r d_ILO}
d_ILO <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=1896457050')
```

```{r d_AoD_from_ILO}
d_AoD_from_ILO <- d_ILO %>%

  rename(ILO = Objectives) %>%

  mutate_if(is.numeric, as.logical) %>%

  # Transform data into tidy format to facilitate manipulation
  gather(key = AoD, value = AoD_covered, `Matrix Complete` : `Intercultural Skills`) %>%

  filter(AoD_covered) %>%

  select(- c(AoD_covered, Comments, Comments_for_Jeroen)) %T>%

  print
```

## Student Data
The tibble `d_transcript` contains the transcript information of students as was provided. It has 40 columns, and rows correspond to a type of grade (e.g. final grade, attendance) per student per course per time they took it.

```{r read_csv_own_parsing}
col_parsing <- cols(
  `Student Number`                      = "c",
  `Program (Abbreviation)`              = "c",
  `Appraisal Status`                    = "c",
  `Module Booking Reason (Description)` = "c",
  `Object name`                         = "c",
  `Start date`                          = "c",
  `End Date`                            = "c"
  )

read_csv_own_parsing <- function(file) file %>% read_csv(col_types = col_parsing)
```

```{r d_transcript, cache = TRUE}
d_transcript <- paste0("Input/Raw Grades/grades", 1:2, ".csv") %>%

  map(read_csv_own_parsing) %>%

  bind_rows

rm(col_parsing, read_csv_own_parsing)
```

# Variable Engineering
## Course Data
The analysis performed on the course data is aimed at discoving what the contribution of each course is towards the fullfilment of the AoDs, and comparing different types of curricula, or programs, based on this infromation.

### AoD
In this analysis, we conside that a course can cover an AoD in two ways: a course covers an AoD if one of its ILOs covers it, or if one of its assessments cover it. For instance, if one of ILO of a course states that the students will learn to analyze empirical data in the context of academic research, then the course in question covers the AoD `Research Skills`; and if one of the assessment is a group presentation, then the course also covers the AoD `Collaborative Skills` and `Communication Skills`.

#### AoD from ILOs
(For this section we use: `d_ILO`-indicates which AoD(s) the intended learning objectives (ILOs) of the courses cover with one row per course-ILO-AoD)

In order to determine which AoD each course covers with its ILOs, we first eliminate the AoD that are not covered by the ILOs, and then we keep only one instance of each combination of course and AoD in case a course had several ILOs covering the same AoD.
```{r d_AoD_ILO}
d_AoD_from_ILO
```

#### AoD from Assessment
(For this section we use: `d_assessment`-indicates which type(s) of assessment each course contains with one row per course-assessment)

In order to determine which AoD each course covers with its assessments, we need to create a binary matrix which indicates which AoD(s) each assessment type covers. We have created such matrix on our google drive and we save it in the following piece of code as `map_assessment_AoD`. `map_assessment_AoD` indicates that, for instance, the assessment type `Essay` covers the AoD `Critical Thinking Skills`, `Communication Skills` and `Writing Skill`. Thus, the first step is to import this matrix:

```{r map_assessment_AoD}
map_assessment_AoD <- gsheet2tbl('https://docs.google.com/spreadsheets/d/1soRA1u5zf9oLNirGmZ9yZ7m2ccPa3XFemnxkj5AVRXo/edit#gid=719531216') %>%

  mutate_if(is.numeric, as.logical) %>%

  # Transform data into tidy format to facilitate manipulation
  gather(key = AoD, value = AoD_covered, `Matrix Complete` : `Intercultural Skills`) %>%

  filter(AoD_covered) %>%

  select(- AoD_covered) %>%

  arrange(Assessment) %T>%

  print
```

Now that we have the matrix `map_assessment_AoD`, we want to find out which AoDs are covered by a course through its assessments. To do this we create an empty tibble `d_AoD_assessment` to store which assessment is covered by each course, and which AoD said assessment covers.Then, we fill in the information with a loop.
In the loop, we first extract a row of `d_assessment` and save it as `observation`. Then, we determine the corresponding assessment which we save as `assessment`. Then, we use the matrix `map_assessment_AoD` to determine which AoD `assessment` covers and use `cbind` and `rbind` to add the information to the tibble `d_AoD_assessment`.
```{r d_AoD_assessment}
d_AoD_from_assessment <- d_assessment %>%

  left_join(map_assessment_AoD, by = "Assessment") %T>%

  print

rm(map_assessment_AoD)
```

#### Combining AoD from ILOs and from assessments
Finally, we can use a `rbind` to combine the two tibbles `d_AoD_ILO` and `d_AoD_assessment`. We also use `distinct` in case a course covers an AoD with both its ILOS and its assessments.
```{r rbind d_AoD_ILO and d_AoD_assessment}
d_AoD <- rbind(
  d_AoD_from_ILO        %>% select(`Course ID`, AoD),
  d_AoD_from_assessment %>% select(`Course ID`, AoD)
  ) %>%

  distinct %>%

  arrange(`Course ID`) %T>%

  print

rm(d_AoD_from_assessment, d_AoD_from_ILO)
```

### Courses
(In this section we use: `d_course`- contains information at the course-level such as in which cluster and concentration(s) they belong, and in which period(s) they are offered.)

Now that we have a clear overview of the distribution of AoDs (`d_AoD`) and assessments (`d_assessment`) among the courses, let us add variables to the tibble `d_course` that contain the information at the course level. For this we create the following three tibbles:
1) `d_ILO_detail` contains two columns indicating the code of the course and the number of ILOs it contains.
2) `d_assessment_detail` contains three columns indicating the code of the course, the number of assessments it covers and a list of the assessments it covers. 
3) `d_AoD_detail` contains three columns indicating the code of the course, the number of AoD it covers and a list of the AoD it covers. 
Then we use a `full_join` to add these variables to the tibble `d_course`.
```{r d_course detail}
count_and_paste <- function(df, var, colnam1, colnam2){

  df %>%

    group_by(`Course ID`) %>%

    summarize(
      !!ensym(colnam1) := n_distinct(!!enquo(var)),
      !!ensym(colnam2) := str_c(!!enquo(var), collapse = ", ")
      )

}

join_by_course_ID <- function(df1, df2) left_join(df1, df2, by = "Course ID")
```

```{r join d_course_detail}
d_course <- d_course %>%

  join_by_course_ID( d_AoD        %>% count_and_paste(AoD       , n_AoD       , `AoD Covered`        ) %T>% print) %>%

  join_by_course_ID( d_assessment %>% count_and_paste(Assessment, n_assessment, `Assessments Covered`)           ) %>%

  join_by_course_ID( d_ILO        %>% count_and_paste(Objectives, n_ILO       , `ILO Covered`        )           ) %>%

  select(
    `Course ID`, `Course Title`, Cluster,
    n_ILO,
    n_assessment, `Assessments Covered`,
    n_AoD, `AoD Covered`,
    everything()
    ) %T>%

  print

rm(count_and_paste, join_by_course_ID)
```

The undegraduate research courses (*UGR-*) are only present at the `3000` level (advanced level). Yet, these course are also offered at the `2000` level (intermediate level). We use an `rbind` to duplicate the rows of the course `UGR3000` and mutate their `Code` to `UGR2000`.
```{r create UGR2000}
UGR_2000 <- d_course %>%

  filter( `Course ID` %in% c("UGR3001", "UGR3002", "UGR3003", "UGR3005")) %>%

  mutate(
    `Course ID` = case_when(
      `Course ID` == "UGR3001" ~ "UGR2001",
      `Course ID` == "UGR3002" ~ "UGR2002",
      `Course ID` == "UGR3003" ~ "UGR2003",
      `Course ID` == "UGR3005" ~ "UGR2005"
      )
    ) %T>%

  print

d_course <- d_course %>% rbind(UGR_2000)

# print UGR course
print(d_course %>% filter(str_detect(`Course ID`, "UGR") ) )

rm(UGR_2000)
```

Finally, we add a series of informative variable at the course level.
```{r d_course additional info}
d_course <- d_course %>%

  separate(
    col     = `Course ID`,
    into    = c("Letters", "Number"),
    sep     = 3,
    remove  = FALSE,
    convert = TRUE
    ) %>%

  mutate(
    Level   = case_when(
      between(Number, 1000, 1999) ~ "Introductory",
      between(Number, 2000, 2999) ~ "Intermediate",
      between(Number, 3000, 3999) ~ "Advanced"    )
    ) %>%

  select(`Course ID`, Letters, Number, Level, everything()) %T>%

  print
```

## Student Data

* Is `Academic Work ID` a unique ID number?

* questions...

(In this section we use: `d_transcript`- contains the transcript information of students as was provided. It has 40 columns, and rows correspond roughly to a type of grade (e.g. final grade, attendance) per student per course per time they took it).

Our dataframe contains many variables and rows that are either empty or meaningless for our analysis. First, we filter, the final grades of studets to keep only relevant rows (Final Confirmed Grades are those with `Appraisal (Description)` as "Grade supervisor"). Then we select ony the 10 variables that we will use for the anlysis, and give them more comprehensible names.

```{r d_transcript helper}
course_letter_UCM <- c("COR", "HUM", "SCI", "SSC", "SKI", "PRO", "UGR", "CAP") %>% paste(collapse = "|")

is_UCM      <- function(course_code) course_code %>% str_detect(course_letter_UCM)
is_elective <- function(course_code) course_code %>% str_detect("HUM|SCI|SSC")
```

```{r d_transcript row selection}
d_transcript <- d_transcript %>%

  # for simplicity, we only keep 
  # *UCM courses*  (courses present in the UCM catalogue)
  # taken in the framework of the *BA Liberal Arts and Sciences (UCM)* (Program (Abbreviation) = 7501)

  filter(
    `Module (Abbrev.)` %>% is_UCM,
    `Program (Abbreviation)` == "7501"
    ) %>%

  # The dataset has multiple rows for each student - course.
  # Following guidelines of Richard Vos, we only consider: 
  # `Appraisal (Description)` == "Grade supervisor"
  # and `Appraisal Type` == "7055"

  filter(
    `Appraisal (Description)` == "Grade supervisor",
    `Appraisal Type`          == "7055" # removes ~ 15 observations
    )
```

```{r d_transcript column selection}
d_transcript <- d_transcript %>%

  # Remove variables with only one value (~10)
  discard(function(x) n_distinct(x) == 1) %>%

  # Select: Student ID, course ID, when a course is taken and the grade 
  select(
    `Student ID`   = `Student Number`,
    `Course ID`    = `Module (Abbrev.)`,
    Year_numerical = `Academic Year`,
    Period         = `Academic Session`,
    Grade          = `Grade symbol`
    )
```

```{r clean_grade}
clean_grade <- function(grade){

  if(is.na(grade) | grade == "NG") grade <- "0"

  grade %>% str_replace(",", ".") %>% as.numeric

}
```

```{r clean_period}
clean_period <- function(period){

  # Course spanning over multiple period are recorder with a 1, 2 or 3
  if(period == 1) return("1 to 6")
  if(period == 2) return("1 to 3")
  if(period == 3) return("4 to 6")

  # Course spanning over a single period are recorded in the 100's, 200's, etc.
  # The digit of the hundred indicates the period.
  period %/% 100 %>% as.character

}
```

```{r clean_year}
clean_year <- function(year) str_c(year, year + 1, sep = "-")
```

```{r clean d_transcript}
d_transcript <- d_transcript %>%

  mutate(

    # Clean grade, period and year
    Grade          = Grade  %>% map_dbl(clean_grade ),
    Period         = Period %>% map_chr(clean_period),
    `Academic Year`= Year_numerical %>% map_chr(clean_year),

    # extract first period in which course spanning over mutliple periods is given
    period_numerical = substr(Period, 1, 1),

    # Combine year and period to have be able to order the course in time using a single variable
    time = str_c(Year_numerical, period_numerical) %>% as.numeric,

    grade_ceil = ceiling(Grade)

    )
```

```{r}
d_transcript_augmented <- d_transcript %>% left_join(select(d_course, - Period), by = c("Course ID"))
```

```{r}
d_transcript_elective <- d_transcript_augmented %>%
  filter(`Course ID` %>% is_elective) %>%
  select(`Student ID`, `Course ID`, time, Grade, grade_ceil)
```

### Issues
```{r issue d_transcript}
d_transcript %>%

  # multiple rows with unique student ID - course ID - time combination.
  add_count(
    `Student ID`, `Course ID`, time,
    sort = TRUE
    ) %>%
  print

d_transcript %>%

  # multiple rows with unique student ID - course ID - time combination - grade.
  add_count(
    `Student ID`, `Course ID`, time, Grade,
    sort = TRUE
    ) %>%
  print

d_transcript %>%

  # SSC2046, a course I took, but excluded from my transcript is absent from the data set
  filter(`Student ID` == "6087587") %>%
  arrange(`Course ID`) %>%
  pull(`Course ID`)

#TODO: list issues here
```

# Save Data
```{r}
save(d_course, list_AoD_assessment, d_AoD, d_assessment, d_ILO, file = "Output/data_general.RDATA")

save(d_transcript_augmented, d_transcript_elective, file = "Output/data_pillar_1.RDATA")

save(d_course, d_transcript, file = "Output/data_pillar_2.RDATA")
```