### 2019-04-02 Pillar 1 - t-test.Rmd

---
title: "Pillar 1 - t-test"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Pillar 1/")

library(tidyverse)
library(arulesSequences)
library(magrittr)

load("Output/data_pillar_1.RDATA")
load("output/d_transcript_cum.RDATA")
```

# Conduct t-test
```{r my_t_test}
d_transcript_cum_informative <- left_join(d_transcript_cum, d_transcript_informative, by = c("sequenceID" = "Student ID", "eventID" = "time"))

my_t_test <- function(predictor, response){

  d_transcript_cum_informative %>%

    filter(`Course ID` == response) %>%

    mutate(has_taken_pred = course_past %>% map_lgl(~ predictor %in% .)) %>% # create two groups: (i) have taken the predictor (ii) have not takne predictor

    t.test(Grade ~ has_taken_pred, data = .)

}
```

```{r d_test}
all_courses <- unique(d_transcript_informative$`Course ID`)

d_test <- crossing(predictor = all_courses, response = all_courses) %>%

  filter(predictor != response) %>%

  mutate(t_test = list(predictor, response) %>% pmap(possibly(my_t_test, otherwise = "only one sample for t test")))
```

# Exploring t-test results

20,000 combinations courses

## Combinations of course for which no t-test was possible ()

8,000 combinations for which one of the sample sizes <= 1.

```{r}
samples <- function(predictor, response){

  d_transcript_cum_informative %>%

    filter(`Course ID` == response) %>%

    summarize(
      n_total   = n(),
      n_taken     = course_past %>% map_lgl(~   predictor %in% .) %>% sum,
      n_not_taken = course_past %>% map_lgl(~ ! predictor %in% .) %>% sum
      )

}

test_impossible <- d_test %>%

  filter(t_test %>% map_lgl(~typeof(.) != "list")) %>%

  mutate(samples = list(predictor, response) %>% pmap(samples)) %>%

  unnest(samples, .drop = TRUE) %>%

  select(response, predictor, everything())
```

## Results of t-test

After filtering, 154 combinations left.

```{r}
test_tidy <- d_test %>%

  filter(t_test %>% map_lgl(~ typeof(.) != "character")) %>%

  mutate(tidied = t_test %>% map(broom::tidy)) %>% # in this case, broom::tidy and broom::glance yield the same results. broom::augment does not work.

  unnest(tidied, .drop = TRUE) %>%

  # disregard advanced-level predictor & introductory-level response
  filter(! str_detect(predictor, "30")) %>% filter(! str_detect(response, "10")) %>%

  # disregard negative estimates & non-significant tests
  filter(estimate > 0) %>% filter(p.value < 0.05) %>%

  group_by(response) %>% arrange(response, desc(estimate)) %>% select(response, predictor, everything())
```