### 2019-04-04 Messy Code/stm single case.Rmd

---
title: "STM single case"
author: "DARS"
date: "4/2/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Setup
##Libraries
```{r library, message = FALSE, eval = TRUE}
library(tidyverse)
library(tidytext)

#Topic Models
library(stm)
library(topicmodels)
library(quanteda) #used for DTM implementation called dfm

library(ggwordcloud) # Word Clouds

library(lemon)
library(ggthemes)
library(rlang) #quote
library(ldatuning) # ideal number of topics in LDA

```

##Loading data
We load the environment `Output/d_text.RDATA`. It contains the data sets:
`d_course`, `d_transcript`, and a list `d_text` with the textual data for catalogues and manuals (for $overview, $description, $manuals we have (`Course ID`, `year`, `word_original`, `word`) ).

For `d_text` we keep only the most recent year.
```{r import data, eval = TRUE}
#load
load("Output/d_text.RDATA")
```

# STM
##Setup
###Cast DTM
The function `my_cast_dtm` casts our dataframes into a DTM (dfm from package quanteda)
```{r my_cast_dtm, eval = TRUE}
my_cast_dtm <- function(data) data %>%
  count(`Course ID`, word) %>%
  cast_dfm(`Course ID`, word, n)

d_cast <- d_text %>% map(my_cast_dtm)

rm(my_cast_dtm)
```

```{r covariates}
#get covariates
covariates <- tribble(
  ~Origin,    ~Data,
  "catalogues", d_text$catalogues %>% distinct(`Course ID`, department),
  "manuals", d_text$manuals %>% mutate(department = "UCM") %>% distinct(`Course ID`, department)
)
```
###STM Corpus
We convert our DTMs into STM corpus to access to full functionality in STM package.
```{r as STM corpus, eval = TRUE}

stm_corpus <- d_cast %>% map(asSTMCorpus)

```

##Fitting STM
We will to train several models to have some flexibility in the number of topics we select. For this purpose we want to create two tibbles. One that trains on the entire data (for use), and one that we divide on training set and test set. 
###Functions
Several functions are needed to work with tibble and tidyverse functionality:
####My_stm Function
The function `my_LDA` fits an LDA model on the DTM using the previous controls.
```{r my_STM, eval= TRUE}
my_stm <- function(n_topics, corpus, data){

  STM_model <- stm(corpus$documents, corpus$vocab, K = n_topics, init.type = "Spectral")
}
```
####My_stm_eval Function
This function applied stm but with heldout vocab and documents. 
```{r}
my_stm_eval <- function(heldout, n_topic){
  stm::stm(heldout$documents, heldout$vocab, n_topic, init.type = "Spectral")
}
```

####Extraction Functions
We will be working with many lists of lists. The following functions help us extract information from lower levels of the list (e.g. access element in second list).
```{r getting gamma and beta, eval = TRUE}
#
##From external Lists:

#Extract DTM
extract_dtm <- function(origin) d_cast[[origin]]
#Extract CORPUS
extract_corpus <- function(origin) stm_corpus[[origin]]

#
## From STM_corpus class

#Extract documents
extract_documents <- function(corpus) corpus$documents
#Extract vocab
extract_vocab     <- function(corpus) corpus$vocab
#Extract data
extract_data     <- function(corpus) corpus$data

#
##From Partition

#Extract heldout docs
extract_held_doc <- function(heldout) heldout$documents#[[1]]
#Extract herlout vocab
extract_held_vocab <- function(heldout) heldout$vocab#[[1]]
#Extract missing
extract_helout_missing <- function(heldout) heldout$missing#[[1]]

#
## MODEL HELDOUT
my_model <- function(heldout){

  docs <- heldout$documents[[1]]
  voc  <- heldout$vocab
  k    <- n_topics

  Model <- stm(docs, voc, 5, init.type = "Spectral")

  return(Model)
}

#
##From Model
#tidy gamma
my_tidy_gamma <- function(Gamma, DTM){
  tidy(Gamma, matrix = "gamma", document_names = rownames(DTM))}

#extract

#extract Distributions (Gamma/Beta)
get_distributions <- function(model, distrib_str){ #input gamma or beta
  tidy(model, matrix = distrib_str)        %>%
    mutate(topic = paste("Topic", topic) ) %>%
      arrange(topic, desc(!!sym(distrib_str)))}

```
####Inspection functions
The following functions allow us to inspect the results from our models(inspect distributions and get kw):
```{r}
get_top_terms_topic <- function(d_beta){
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(10, beta) %>%
    select(topic, term) %>%
    group_by(topic) %>%
    summarise(top_words = list(term))

}

get_top_topics_course <- function(d_gamma){
  data.frame(d_gamma) %>%
    group_by(document) %>%
    top_n(5, gamma) %>%
    select(document, topic) %>%
    summarise(top_topic = list(topic))
}

get_kw <- function(d_beta){
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(3, beta) %>%
    ungroup()%>%
    pull(term)%>%
    unique()
}
```
###Model Tibbles
####For use:
```{r models in tibble}
topic_numbers <- seq(from = 5, to = 10, by = 5) # seq(from = 5, to = 20, by = 5) #seq(from = 5, to = 150, by = 5)

stm_model <- tibble(
  #repeat each topic number twice (once for overview and once for manual)
  n_topic = rep(topic_numbers, each = 2) ) %>%
  mutate(
  #Identification information
  Origin     = rep(c("catalogues", "manuals"), length(topic_numbers)),         
  `Topic ID` = paste(Origin, n_topic, sep = "_"),
  #DTM

  DTM       = pmap(.l = list(Origin), .f = extract_dtm),

  doc_names = map(DTM, rownames),
  #STM Corpus
  Corpus           = pmap(.l = list(Origin),   .f = extract_corpus), 
  #Documents- STM corpus
  Documents_stm     = pmap(.l=list(Corpus),    .f = extract_documents),
  #Vocab- STM corpus
  Vocab_stm         = pmap(.l=list(Corpus),    .f = extract_vocab),
  #Data- STM corpus
  Data_stm          = pmap(.l=list(Corpus),    .f = extract_data),
  #Fit Model
  Model  = pmap(.l = list(n_topic, Corpus), .f = my_stm),

  #Evaluation
  Exclusivity        = map(Model, exclusivity),

  semantic_coherence = pmap(.l = list(Model,DTM), semanticCoherence, M = 10), #M = num. of top words per topic
  #Alretnative input to semantic_coherence = pmap(.l = list(Model, Documents_stm), semanticCoherence, M = 10),

  #
  ##Inspection
  #Beta distribution
  Beta = pmap(.l = list(Model),  .f = tidy),

  #Gamma distribution
  Gamma = pmap(.l = list(Model, DTM), .f = my_tidy_gamma), #fix this

  #top words per topic
  top_terms_topic = map(Beta, .f = get_top_terms_topic),

  #top topics per course
  top_topic_course = map(Gamma, .f = get_top_topics_course),

  #Select keywords
  kw = map(Beta, .f = get_kw)

)

```
####TESTING AND INSPECTING MODELS
```{r model inspection}
my_test <-  stm_model %>% filter(n_topic == 10, Origin == "manuals")
```
####For evaluation
```{r models in tibble}
topic_numbers <-  seq(from = 5, to = 10, by = 5) #seq(from = 5, to = 150, by = 5)

#
##STM EVALUATION TIBBLE
stm_eval_model <- tibble(

  #repeat each topic number twice (once for overview and once for manual)
  n_topic = rep(topic_numbers, each = 2)

  ) %>%
  mutate(
    #Identification information
    Origin     = rep(c("overview", "manuals"), length(topic_numbers)),         
    `Topic ID` = paste(Origin, n_topic, sep = "_"),

    #DTM
    DTM       = pmap(.l = list(Origin),  .f = extract_dtm),
    Heldout   = pmap(.l = list(DTM),     .f = make.heldout, proportion = 0.5),

    #Fitting Model
    Model     = pmap(.l = list(Heldout, n_topic), .f = my_stm_eval),

    #Evaluation measures
    Exclusivity        = map(Model, exclusivity),

    semantic_coherence = pmap(.l = list(Model, Heldout), 
                              function(model,  heldout) semanticCoherence(model, heldout$documents)),

    eval_heldout       = pmap(.l = list(Model, Heldout), 
                              function(model,  heldout) eval.heldout(model, heldout$missing)),

    residual           = pmap(.l = list(Model, Heldout), 
                              function(model,  heldout) checkResiduals(model, heldout$documents, tol = 1/100)),

    bound              = pmap_dbl(.l = list(Model), 
                                  function(model) max(model$convergence$bound)),

    lfact              = map_dbl(Model,       
                                 function(model) lfactorial(model$settings$dim$K)),

    lbound             = bound + lfact,
    iterations         = map_dbl(Model,       
                                 function(model) length(model$convergence$bound))
    ) 
```