### 2019-04-15 Data Preparation - Course Catalogues.Rmd
Adds MSP catalogue

---
title: "Data Preparation - Course Catalogues"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Data Preparation/")

library(tidyverse)
library(tidytext)
library(gsheet) # import spreadsheets from google drive
library(tm)
library(hunspell) # Stemmer
library(magrittr)
```

# Helper functions

```{r helpers}
extract_id_pdf    <- function(string) string %>% str_sub(start = 1, end = - 5)
extract_id_course <- function(string) string %>% str_sub(start = 1, end =   7)

collapse_or       <- function(string) string %>% str_c(collapse = "|")
collapse_space    <- function(string) string %>% str_c(collapse = " ")

symbol_to_line    <- c("\r\n", "\n") %>% collapse_or
split             <- function(string) string %>% str_split(pattern = symbol_to_line) %>% .[[1]]

remove_header_UCV <- function(string) string %>% str_sub(28, -1)

overview_section_start <- function(string) string %>% str_detect(pattern = "^ *Core Courses|^ *Mathematics and")
overview_section_end   <- function(string) string %>% str_detect(pattern = "Appendix|^ *Courses at\n *Maastricht Science Programme|^*Courses Available at University")

headers_UCM_UCV_MSP <- c("^ *Core Courses",
                     "^Humanities \\(HUM\\)",
                     "^ *Life Sciences Courses",
                     "^Sciences \\(SCI\\)",
                     "^ *Social Sciences",
                     "^ *Skills Trainings",
                     "^ *Projects",
                     "^Undergraduate Research",
                     "Appendix",
                     "^ *Mathematics and") %>% collapse_or
is_header     <- function(string) string %>% str_detect(pattern = headers_UCM_UCV_MSP)

course_code_UCM_UCV_MSP <- c("COR", "HUM", "SCI", "SSC", "SKI", "PRO", "UGR", "CAP", # UCM
                         "VCO", "VSC", "VSS", "VSK", "VPR", "VCA", # UCV
                         "BIO", "CHE", "MAT", "PHY", "NEU", "INT", "PRA" #MSP
                         ) %>% collapse_or
overview_start  <- function(string) string %>% extract_id_course %>% str_detect(pattern = course_code_UCM_UCV_MSP) # simplify with %in% course_code_UCM
```

# Importing course data

```{r read_in_pdfs}
read_in_pdfs <- function(file){

  Corpus(
    x             = DirSource(file),
    readerControl = list(reader = readPDF(control = list(text = "-layout")))
    ) %>%

    # turn corpus into a tibble to facilitate manipulation
    as.list %>% tibble(raw = .) %>%

    transmute(
      ID   = raw %>% map(function(x) x$meta$id) %>% map_chr(extract_id_pdf),
      text = raw %>% map(content) 
      )

}
```

## Application
```{r d_text}
d_text <- list()
```

```{r d_text$catalogues}
d_text$catalogues <- tribble(
  ~ ID              ,  ~ page_top                  , ~ before_overview                       ,
  "European_studies", "^.Bachelor European Studies", "Faculty of Arts and Social Sciences"   ,
  "Psychology"      , "^.Bachelor Psychology"      , "Faculty of Psychology and Neuroscience",
  "UCV"             , NA_character_                , NA_character_                           ,
  "UCM"             , NA_character_                , NA_character_                           ,
  "MSP"             , NA_character_                , NA_character_
  ) %>%

  full_join(read_in_pdfs("./Input/Catalogues/UM"), by = "ID") %>%

  mutate(year = "2018-2019") %>%
  rename(catalogue = text)
```

```{r d_text$manuals}
d_text$manuals <- read_in_pdfs("./Input/Manuals 2017-18") %>% rename(`Course ID` = ID) %>% mutate(text = text %>% map_chr(str_c, collapse = " "))
```

# Extracting course overviews

In order to conduct a topic analysis of course content, we extract the *overview* and *description* of each course from the course catalogues, as well as the  *full text*  from the Course Manuals.

(For this section we use:`corpus_catalogues` contains the pdfs of the `5` most recent course catalogues.)

We start by extracting the overviews which are 1 or 2 page long and contain important information of a course. To accomplish this, for each catalogue (loop), we first use `grep` to determine which pages should be excluded from the analysis (`pages_to_exclude`). These pages are typically headers. Excluding them before the analysis allows use to keep the code relatively simple. We also make use the fact that the overviews always start with the code of the course, i.e. it starts with one of the elements of `course_code`; and we use `grep` to identify the first page of each overview. We use grep a 3rd time to identify the page containing the header *Core Coureses (COR)* and which marks the beginning of the overviews in the catalogue, and we loop from this page to the last page of the catalogue.

In the loop, we first determine the content of the current and following page. If the current page is the first page of an overview, then we identify the overview (one or two pages) and save it to the tibble `d_description`. To identify the overview of a course, we check if the following page is either the first page of an overview or a page to exclude. If it is either of these, then the overview correspond to the content of the current page; but if the following page is neiter of these, then the overview correspond to the content of the current page and that of the following page.

## ES, PSY
```{r overview_ES_PSY}
overview_ES_PSY <- function(catalogue, ID, page_top, before_overview){

  catalogue %>%

    # create tibble to facilitate manipulation
    collapse_space %>% split %>% tibble(line = .) %>%

    # remove page numbers and page top to faciliate the identification of the course description starts
    mutate(page_top    = str_detect(line, page_top),
           page_number = page_top %>% lead(default = FALSE)) %>%

    filter(!page_top, !page_number) %>%

    # identify course descriptions
    mutate(overview_start = str_detect(line, before_overview) %>% lag(default = FALSE)) %>%
    mutate(overview_id    = cumsum(overview_start)) %>%

    filter(overview_id != 0) %>%

    # extract course description
    group_by(overview_id) %>% 
    summarize(text        = line %>% str_c(collapse = " "),
              `Course ID` = line %>% str_c %>% .[[1]]) %>%
    mutate(department  = ID) %>%
    select(-overview_id)

}
```

## UCM, UCV, MSP

```{r paste_if_2_page_overview}
paste_if_2_page_overview <- function(page, page_following){

  if(! page %>% overview_start) NA_character_ # if the page is not the start of an overview, return an NA

  else{
    if(page_following %>% overview_start | is.na(page_following))  page # include possibility of NA for last page of overview section
    else                                                           str_c(page, page_following)
  }

}
```

```{r overview_UCM}
overview_UCM_UCV_MSP <- function(catalogue, ID){

  if(ID == "UCV") catalogue <- catalogue %>% remove_header_UCV

  # create tibble to facilitate manipulation
  catalogue %>% tibble(page = .) %>%

  # identify start and end of overview section of catalogue
  mutate(overview_section_start = page %>% overview_section_start,
         overview_section_end   = page %>% overview_section_end  ) %>%

  # filter the overview section
  mutate(after_overview_section_start = cumany( overview_section_start),
         before_overview_section_end  = cumall(!overview_section_end  )) %>%
  filter(after_overview_section_start & before_overview_section_end    ) %>%

  # exclude headers to only keep the course overviews themselves
  filter(! page %>% is_header) %>%

  # paste current page and following page (`lead(page)`) if overview is two page long
  transmute(
    text        = list(page, lead(page)) %>% pmap_chr(paste_if_2_page_overview),
    `Course ID` = text %>% extract_id_course) %>%
    mutate(department  = ID) %>%
    filter(!is.na(text))

}
```

## Application
```{r overview}
overview <- function(ID, catalogue, page_top, before_overview){

  if(ID %in% c("European_studies", "Psychology")){
    catalogue %>% overview_ES_PSY(ID = ID, page_top = page_top, before_overview = before_overview)

  }else if(ID %in% c("UCM", "UCV", "MSP")){
    catalogue %>% overview_UCM_UCV_MSP(ID = ID)

  } else "to do"

}
```

```{r}
d_text$catalogues <- d_text$catalogues %>% mutate(overviews = list(ID, catalogue, page_top, before_overview) %>% pmap(overview))
```

```{r}
d_text$catalogues <- d_text$catalogues$overviews %>% bind_rows
```

# Clean data

## Tidy data
To put everything into tidy text format, we first create three tibbles `d_overview_tidy`, `d_description_tidy`, `d_manual` that respectively store the overviews, descriptions and text from manuals in the tidy text format, with one row per course-year-word (and course-word for `d_manual`).

```{r}
tidy_text <- function(df) df %>% unnest_tokens(output = word, input = text)
```

```{r tidy d_text}
d_text <- d_text %>% map(tidy_text) %T>% print

rm(tidy_text)
```

## Stem data
We then use the function `hunspell_stem`, which returns valid stems for a given word, to stem the words in the tibbles `d_overview_tidy` and `d_description_tidy`.

We first create a function `stem_hunspell` which, given a term, returns its most basic stem (the last one from the list of stem returned by `hunspell_stem`). We would like to apply `stem_hunspell` on the words of `d_overview_tidy` and `d_manual`, since they have similar structures, we row bind them into `dictionary` to ease the application of `stem_hunspell`. However, since, `stem_hunspell` is not a vectorized function and the number of words in `dictionary` is large, we first use distinct on the `word` variable to find a list containing *once* each term present in the course overviews and manuals. Then, we use `mutate` to apply the function `stem_hunspell` on each word from our dictionary and save its results as a new column `word_stem`.

```{r stem}
stem <- function(term) {

  stems   <- hunspell_stem(term)[[1]] # look up the term in the dictionary
  n_stems <- length(stems)            
  if (n_stems == 0) term              # if no stems, return original term
  else              stems[[n_stems]]  # if multiple stems, return last (most basic) stem

}
```

```{r dictionary, cache = TRUE}
dictionary <- union(d_text$catalogues$word, d_text$manuals$word) %>%

  tibble(word = .) %>%

  mutate(word_stem = word %>% map_chr(stem))

# terms for which the stem differs from the original word
filter(dictionary, word != word_stem)
```

```{r destem}
# TODO: "destem" terms by hand: cuss => discuss
filter(dictionary, word_stem == "cuss")
```

Finally, we create a function `stem_with_dictionary` that performs a left join on the dataframe it takes as input with `dictionary`, thus adding the word stems to the original dataframe.

Then, we use `stem_with_dictionary` to include the stems of all words in `d_description`, `d_overview`, and `d_manual`.

```{r stem_with_dictionary}
stem_with_dictionary <- function(data){

  data %>% left_join(dictionary, by = "word") %>%

    rename(word_original = word,
           word          = word_stem)

}
```

```{r stem d_text}
d_text <- d_text %>% map(stem_with_dictionary)

rm(stem, stem_with_dictionary)
```

# Remove stopwords

Finally, we want to filter out some uninformative words from our textual data. For this, we first store all the unwanted words in a vector `sw_owm`. We also create function `remove_sw` which receives a dataframe as input and filters out all words in `sw_own`. Then we apply this function to `d_description`, `d_overview`, and `d_manual`.
```{r sw_own}
sw_own <- 1 : 1e3 %>% as.character %>% tibble( word = .)
```

```{r remove_sw}
remove_sw <- function(data){

  data %>%

    anti_join(stop_words, by = "word") %>%
    anti_join(sw_own    , by = "word") %>%

    # remove words of 1 or 2 letters
    filter(nchar(word) > 2) %>%

    # Remove words occuring once or twice in corpus (usually names, website links or typos)
    add_count(word) %>%
    filter(n > 2) %>%
    select(-n) %>%

    # remove NA
    filter(!is.na(word))

}
```

```{r sw d_text}
d_text <- d_text %>% map(remove_sw) %T>% print

rm(sw_own, remove_sw)
```

# Current course
```{r}
course_current <- d_text$catalogues %>% filter(department == "UCM") %>% distinct(`Course ID`) %>% pull
```

# Save data

```{r}
save(course_current, file = "Output/course_current.RDATA")
save(d_text        , file = "Output/d_text.RDATA")
```