### 2019-04-17 Miscellaneous code /Experiments/Ternary Plots.Rmd
moved things to different folder

---
title: "Ternary plot"
author: "DARS"
date: "4/17/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Ternary Plots
This is a simple visualization tool to see where you are in the SCI, SSC, and HUM space.

```{r libraries}
library(tidyverse)
library(Ternary)

```

## Load
```{r load, echo=FALSE}
load("~/Documents/DARS/GitDARS/Output/data_pillar_1.RDATA")
```

```{r}
data <- d_transcript_augmented %>% 
  select(`Student ID`, `Course ID`) %>% 
  distinct() %>%
  mutate(`Course Concentration` = str_sub(`Course ID`, 1,3)) %>%
  filter(`Course Concentration` %in% c("HUM", "SSC", "SCI")) %>%
  group_by(`Student ID`) %>%
  count(`Course Concentration`) %>%
  group_by(`Student ID`) %>%
  mutate(student_total = sum(n),
         proportion = n/student_total) %>%
  select(-n,-student_total) %>%
  spread(key = `Course Concentration`, value = proportion, fill = 0)

```

```{r}
TernaryPlot()
AddToTernary(points, data)

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.