### 2019-04-23 Pillar 2 - Tmp Topic Modeling.Rmd
evaluate over many models catalogues

---
title: "Pillar 2- Quick Topic Models"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Pillar 2/",
                      eval = FALSE)
```

#Setup
##Libraries
```{r library, message = FALSE, eval = TRUE}
library(tidyverse)
library(tidytext)

library(topicmodels)
library(lemon)
library(ggthemes)
library(rlang) #quote
library(ldatuning) # ideal number of topics in LDA

```
##Loading data
We load a list `d_text` with two tibbles of textual data for catalogues and manuals. That is, for $catalogues and $manuals we have a tibble with the following columns: `Course ID`, `year`, `word_original`, `word`.

We also load the environment `Output/data_pillar_2.RDATA` containing the data sets`d_course` and `d_transcript` that we will need to prepare some imputs for our App.

```{r import data, eval = TRUE}
#load
load("./Output/d_text.RDATA")
load("./Output/data_pillar_2.RDATA") #used for app

```
###Preparation of course_title vector
```{r course titles, eval = TRUE}
#Clean titles
course_titles <-  d_text$catalogues %>%
  select(`Course ID`, `Course Title`, department) %>% 
  distinct

empty_titles <- course_titles %>% filter(`Course Title` == "")

```
###Functions for internal faceting
We create the functions for internal faceting taken from https://github.com/dgrtwo/drlib/blob/master/R/reorder_within.R
```{r function reorder_within & co., echo = FALSE, eval = TRUE}

# Function for ordering within facet:
# https://github.com/dgrtwo/drlib/blob/master/R/reorder_within.R

reorder_within <- function(x, by, within, fun = mean, sep = "___", ...) {
  new_x <- paste(x, within, sep = sep)
  stats::reorder(new_x, by, FUN = fun)
}

scale_x_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_x_discrete(labels = function(x) gsub(reg, "", x), ...)
}

scale_y_reordered <- function(..., sep = "___") {
  reg <- paste0(sep, ".+$")
  ggplot2::scale_y_discrete(labels = function(x) gsub(reg, "", x), ...)
}
```

# LDA
In this section, we train several LDA models on our data and evaluate these to keep the best topic model for the course content data we have. We will do the model training and evaluation measures in one go using mapping functions on a tibble containing all pertinent data. In this way we ensure that we maintain the order of all the information we gather on our models.

In order to do this, we have to create all our functions before hand, to apply them in a chain later. Moreover, we will need to create a document-term matrix (DTM) for the catalogues and manuals which we will use in several steps to create and evaluate the model. Thus, we will create the DTM in a separate step as part of the set up.

###Cast DTM
The function `my_cast_dtm` casts each of our dataframes (catalogues and manuals) into a DTM.
```{r my_cast_dtm, eval = TRUE}
#
##Function-----------------------------------------------
my_cast_dtm <- function(data) data %>%
  count(`Course ID`, word) %>%
  cast_dtm(`Course ID`, word, n)

#
##Application-----------------------------------------------
d_cast <- d_text %>% map(my_cast_dtm)

#
##Clean workspace-----------------------------------------------
rm(my_cast_dtm)

```
###Controls
The control parameters we will use to train our LDA model are:
```{r controls, eval = TRUE}
my_control <- list(

  nstart = 10,
  seed   = 1 : 10, #must have length nstart
  best   = TRUE,

  burnin = 100,
  iter   = 2000,
  thin   = 2000

)
```
We create a helping control that we can use when we need fast convergence for testing new code. Thus, the control parameters we use for fast testing are:
```{r fast control, eval = TRUE}
my_control_fast <- list(

  nstart = 1,
  seed   = 1,
  best   = TRUE,

  burnin = 5,
  iter   = 10,
  thin   = 2

)
```

## Fitting Model
The function `my_LDA` fits an LDA model on the DTM using the previous controls.
```{r my_LDA, eval= TRUE}
my_LDA <- function(n_topics, corpus, control = my_control){

  LDA_model <- LDA(
    x       = corpus,
    k       = n_topics,
    method  = "Gibbs",
    control = control
    )
}

```

###Extracting functions
The following functions are used as input in a map to compute and evaluate topic models. 
```{r getting gamma and beta, eval = TRUE}
#Extract DTM-----------------------------------------------
extract_dtm <- function(origin) d_cast[[origin]]

#Extract Distributions (Gamma/Beta)-----------------------------------------------
get_distributions <- function(model, distrib_str) { #input gamma or beta
  tidy(model, matrix = distrib_str) %>%
    mutate(topic = paste("Topic", topic) ) %>%
    arrange(topic, desc(!!sym(distrib_str)))
}

#perplexity Function-----------------------------------------------
my_perplexity <- function(model, dtm) {
  perplexity(model, dtm, estimate_theta = FALSE)
}

```

###Inspection functions
The following functions allow us to inspect the results from our models(inspect distributions and get kw):
```{r inspection functions, eval = TRUE}

get_top_terms_topic <- function(d_beta) {
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(10, beta) %>%
    select(topic, term) %>%
    group_by(topic) %>%
    summarise(top_words = list(term))

}

get_top_topics_course <- function(d_gamma) {
  data.frame(d_gamma) %>%
    group_by(document) %>%
    top_n(5, gamma) %>%
    select(document, topic) %>%
    summarise(top_topic = list(topic))
}

get_kw <- function(d_beta) {
  data.frame(d_beta) %>%
    group_by(topic) %>%
    top_n(3, beta) %>%
    ungroup() %>%
    pull(term) %>%
    unique()
}

```
###Training Models
We train the LDA for different numbers of topics (5 to 250 by 5), we save the resulting 50 models in a tibble under the name `tb_topic_model`, along with the DTM used to build the model, the corpus of origin (catalogue or manuals), the word and topic distributions results from that model, it's evaluation measures, and vectors of words and topics that allow us humans to get a feel of the model, along with some key words we will use in the app.
This step takes some time, thus we also measure how long it takes to train the model.
```{r LDA fit full, cache = TRUE, eval = FALSE}
#Timing_____________________________________________________________________
initial_time <- Sys.time()

#Training model______________________________________________________________
topic_numbers <-c(10, 25, 30, 35, 40, 45, 50, 55, 60) #seq(from = 5, to = 10, by = 2) #5 to 150 by 50

tb_topic_model <- tibble(

  #repeat each topic number twice (once for catalogues and once for manual)
  n_topic = rep(topic_numbers, each = length(d_text))

  ) %>%
  mutate(Origin     = rep(c("catalogues", "manuals"), length(topic_numbers)),         # should be 50
         `Topic ID` = paste(Origin, n_topic, sep = "_"),

         #Document Term Matrix
         DTM        = pmap(.l = list(Origin), .f = extract_dtm),

         #Fit Model
         Model      = pmap(.l = list(n_topic, DTM), .f = my_LDA, control = my_control), #my_control

         #Gamma distribution
         Gamma      = pmap(.l = list(Model), .f = get_distributions, distrib_str = "gamma"),

         #Beta distribution
         Beta       = pmap(.l = list(Model),  .f = get_distributions, distrib_str = "beta"),

         #Perplexity
         Perplexity = pmap_dbl(.l = list(Model, DTM), .f = my_perplexity),

         #Loglikelihood
         LogLikelihood = pmap_dbl(.l = list(Model), .f = logLik), #this removes the degrees of freedom data

         #
         ##Inspect

         #top words per topic
         top_terms_topic = map(Beta, .f = get_top_terms_topic),

         #top topics per course
         top_topic_course = map(Gamma, .f = get_top_topics_course),

         #Select keywords
         kw = map(Beta, .f = get_kw)
         )

#Timing______________________________________________________________
final_time <- Sys.time()
took_how_long <- final_time - initial_time

#View tibble______________________________________________________________
tb_topic_model

#Saving______________________________________________________________
#save(tb_topic_model, "tmp_topic_model_tibble.RDATA")#DOESNT WORK

# Clean workspace
# rm(extract_dtm, get_distributions, my_perplexity,
#    get_top_terms_topic, get_top_topics_course, get_kw)

```
###Training Models separately
For time constrains, we sometimes need to train the model only on the manuals or the catalogues, for this we add an extra filtering step in the code that filters on `which_texts`.
```{r LDA fit separate, cache = TRUE, eval = TRUE}
#Timing
initial_time <- Sys.time()

#Training
topic_numbers <-seq(from = 5, to = 150, by = 5)#c(50, 55, 60, 65, 70) #seq(from = 5, to = 10, by = 2) #5 to 150 by 50
which_texts <- "catalogues"

tb_topic_model <- tibble( 
  n_topic = rep(topic_numbers, each = length(d_text)) 
  ) %>%

  mutate(Origin     = rep(c("catalogues", "manuals"), length(topic_numbers)),         # should be 50
         `Topic ID` = paste(Origin, n_topic, sep = "_"),

         #Document Term Matrix
         DTM        = pmap(.l = list(Origin), .f = extract_dtm)) %>%

  filter(Origin == which_texts) %>%

  mutate(
         #Fit Model
         Model  = pmap(.l = list(n_topic, DTM), .f = my_LDA, control = my_control), #my_control

         #Gamma distribution
         Gamma = pmap(.l = list(Model), .f = get_distributions, distrib_str = "gamma"),

         #Beta distribution
         Beta = pmap(.l = list(Model),  .f = get_distributions, distrib_str = "beta"),

         #Perplexity
         Perplexity = pmap_dbl(.l = list(Model, DTM), .f = my_perplexity),

         #Loglikelihood
         LogLikelihood = pmap_dbl(.l = list(Model), .f = logLik), #this removes the degrees of freedom data

         #
         ##Inspect

         #top words per topic
         top_terms_topic = map(Beta, .f = get_top_terms_topic),

         #top topics per course
         top_topic_course = map(Gamma, .f = get_top_topics_course),

         #Select keywords
         kw = map(Beta, .f = get_kw)
         )

#Timing
final_time <- Sys.time()
took_how_long <- final_time - initial_time

#View
#tb_topic_model

#Save
save(tb_topic_model, file = paste("./Output/", which_texts, "tmp_topic_model_tibble.RDATA", sep = "_"))

```

##Plots - Model Evaluation
To evaluate how many topics are optimal, we graph the perplexity and log Likelihood of models against the number of topics.
```{r perplexity plots, eval = FALSE}
#Function
plot_evaluation <- function(origin) {

  #data
  tb_topic_model %>%

    #Wrangling
    select(n_topic, Origin, Perplexity, LogLikelihood) %>%
    gather(key = "Eval Type", value = "Eval Value", c(Perplexity, LogLikelihood)) %>%
    filter(Origin == origin) %>%

    #Graphics
    ggplot(mapping = aes(x = n_topic, y = `Eval Value`)) +
    geom_point() +
    geom_line() +
    facet_grid(rows = vars(`Eval Type`), scales = "free") +
    theme_minimal()

}

#Plot
plot_evaluation("catalogues")
#plot_evaluation("manuals")
```
##Different alphas and delta (prior of term distributions over topics)
To evaluate the impact of the initial alpha and beta on our model, we train the model using different inizialization values for alpha and delta. Note that this is only an inizialisation and the algorithm will try to extimate alpha and beta.
###Alpha
```{r alpha, eval = FALSE}
sample_dtm <- tb_topic_model$DTM[[1]] #1 CATALOGUES, 2 MANUALS

fit_lda_alpha <-  function(alpha) {
  control = list(
        nstart = 10,
        seed   = 1:10,
        best   = TRUE,
        burnin = 5,
        iter   = 200, #2000
        thin   = 200, #2000
       alpha = alpha)

  topicmodels::LDA(x = sample_dtm, k = 50, method = "Gibbs", control = control)

}
# test
fit_lda(0.1)
#Evaluation
tm_eval_alpha <- tibble( alpha = seq(from = 0.001, to = 2, length = 20)) %>%
  mutate(Model = map(alpha, .f =fit_lda_alpha),
         DTM = list((testing_dtm)),
         Gamma = pmap(.l = list(Model), .f = get_distributions, distrib_str = "gamma"),

         #Beta distribution
         Beta = pmap(.l = list(Model),  .f = get_distributions, distrib_str = "beta"),
         #Perplexity
         Perplexity = pmap_dbl(.l = list(Model, DTM), .f = my_perplexity),

         #Loglikelihood
         LogLikelihood = pmap_dbl(.l = list(Model), .f = logLik))
```
#Beta
```{r beta, eval = FALSE}

fit_lda_delta <-  function(delta) {
  control = list(
        nstart = 10,
        seed   = 1:10,
        best   = TRUE,
        burnin = 5,
        iter   = 200, #2000
        thin   = 200, #2000
        delta  = delta)

  topicmodels::LDA(x = sample_dtm, k = 50, method = "Gibbs", control = control)

}
# test
fit_lda_delta(0.1)

#Evaluation
tm_eval_delta <- tibble( delta = seq(from = 0.001, to = 2, length = 20)) %>%
  mutate(Model = map(delta, .f =fit_lda_delta),
         DTM = list((testing_dtm)),
         Gamma = pmap(.l = list(Model), .f = get_distributions, distrib_str = "gamma"),

         #Beta distribution
         Beta = pmap(.l = list(Model),  .f = get_distributions, distrib_str = "beta"),
         #Perplexity
         Perplexity = pmap_dbl(.l = list(Model, DTM), .f = my_perplexity),

         #Loglikelihood
         LogLikelihood = pmap_dbl(.l = list(Model), .f = logLik))
```

##Saving model_data:
We save the data from our intermediate step in case further analysis is needed.
```{r save, eval = TRUE}
#save(tb_topic_model, file = "Output/tb_topic_model.RDATA" )
```

#App
For our Recommender System, we need to pass the distributions, number of topics, origin, and key words to the app.
##Distributions and Key words
The function `clean_kw` removes stopwords from our key words.
```{r function clean key words, eval = T}
clean_kw <- function(kw_list) {

  #stopwords
  kw_stopwords <- c("source", "primary", "type", "assess", "involve", "e.g", "discussion", "institute", "ha", "introduce",
                 "role","address", "student", "examination","coordinator", "maastrichtuniversity.nl", "identify", "manner", "mechanism", "concept", "hour",
                 "teach",  "depth","question", "specific", "busy","academic", "semester", "week", "test", "session", "team", "communicate", "theory", "text",
                 "library", "article", "examine", "current", "routledge", "temporary", "flow", "level", "study", "task")

  #remove stopwords
  kw <- setdiff(kw_list, kw_stopwords)
  #return
  return(kw[!is.na(kw)])

}

```
On our `topic_model` tibble we filter only information relevant for the app, and apply `clean_kw` on our `kw` column
```{r crop and clean TM for App, eval = T}
app_topic_model <- tb_topic_model %>% 
  select(`Topic ID`, n_topic, Origin, Gamma, Beta, kw) %>%
  mutate(kw = map(kw, .f = clean_kw))
```

##Convenience Vectors
For the app we also need to have a list of all courses and an example of a prospective semester
```{r for convenience, eval = T}
# set seed
set.seed(1)

#
## All courses
course_all <- d_course %>%
  inner_join(
    d_transcript,
    by = "Course ID"
  ) %>%
  distinct(
    `Course ID`
  ) %>%
  # remove semester abroad, skills and projects
  filter(
    ! str_detect(`Course ID`, pattern = "SA|SKI|PRO")
  ) %>%
  pull

#
##courses following_semester
course_following_semester <- sample(x = course_all,
                                    size    = 60,
                                    replace = FALSE) %>% 
  sort()

```

```{r adding convenience vectors, eval = T}
app_topic_model <- app_topic_model %>%
  mutate(`All Courses`     = list(course_all),
         `Sample semester` = list(course_following_semester),
         `Course Titles`   = list(course_titles) )

app_model <- app_topic_model %>%
  filter(n_topic == 55)

#rm(course_all, course_following_semester)
```
## Saving files for APP
We save all requirements for the app.
```{r app files, eval = T}
save(app_topic_model, app_model, file = "App/Recommender System/data_topic_models.RDATA")
```