### 2019-04-23 Miscellaneous code /Experiments/Read in articles for recommendation.Rmd
adds files to input and output materials

---
title: "Read in articles for recommendation"
author: "DARS"
date: "4/18/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
##Libraries
```{r library, message = FALSE, eval = TRUE}
library(tidyverse)
library(tidytext)

library(topicmodels)
library(lemon)
library(ggthemes)
library(rlang) #quote
library(ldatuning) # ideal number of topics in LDA

library(tm)
library(hunspell) # Stemmer
library(magrittr)

```
#Load set up
```{r}
load("~/Documents/DARS/GitDARS/App/Recommender System/data_topic_models.RDATA")
load("~/Documents/DARS/GitDARS/Output/_catalogues_tmp_topic_model_tibble.RDATA")
```

We select a single model to pass it the new document
```{r load model}
my_model <- tb_topic_model %>% filter(n_topic == 55) %>% .[["Model"]]
```

We load a new document and clean it
```{r}

read_in_pdfs <- function(file) {

  Corpus(
    x             = DirSource(file),
    readerControl = list(reader = readPDF(control = list(text = "-layout")))
    ) %>%

    # turn corpus into a tibble to facilitate manipulation
    as.list %>% tibble(raw = .) %>%

    transmute(
      ID   = raw %>% map(function(x) x$meta$id),
      text = raw %>% map(content)
      )

}
```

```{r}
my_pdf <- read_in_pdfs("~/Documents/DARS/GitDARS/Input/Articles/")

my_pdf$text

my_document <- tibble( 
  ID   = my_pdf$`Narrative Media-Article.pdf`$meta$id,
  page = my_pdf$`Narrative Media-Article.pdf`$content)

my_document <- my_document %>% 
  mutate(ID = str_sub(ID, 1,-5)) %>% #remove .pdf
```