### 2019-04-30 Pillar 1 - Topic Profile.Rmd

---
title: "Pillar 1 - Student Topic Profile"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

__Considerations__:

* extract course descriptions from courses not offer in 2018-2019 e.g. SCI2012.
* give more weight to 3000-level courses

* consider following predictors:
     * GPA (per concentration)

```{r setup, include = FALSE}
knitr::opts_chunk$set(cache.path = "Cache/Pillar 1/")

library(magrittr)
library(glmnet)
library(leaps)
library(tidyverse)
```

```{r load, cache.lazy = TRUE}
#load("App/Recommender System/topic_model_gb.RDATA")
load("App/Recommender System/data_topic_models.RDATA")
load("Output/data_pillar_1.RDATA")
load("Output/course_current.RDATA")

course_target <- course_current[! str_detect(course_current, "...10..")]
course_target <- union(course_target, str_subset(course_current, "COR"))# include core courses
```

# Course topic profile
```{r gamma_spread}
gamma        <- app_model$Gamma %>% .[[1]] %>% mutate(topic = topic %>% str_replace(" ", "_"))  #topic_model_gb$g_overview[[15]] %>% mutate(topic = topic %>% str_replace(" ", "_"))
gamma_spread <- gamma %>% spread(topic, gamma)

beta <- app_model$Beta %>% .[[1]] %>% group_by(topic) %>% top_n(15, beta) #topic_model_gb$b_overview[[15]] %>% group_by(topic) %>% top_n(15, beta)
```

# Student topic profile
```{r student_TP}
student_TP <- d_transcript_augmented %>%

  left_join(gamma_spread, by = c("Course ID" = "document")) %>%

  mutate_at(vars(matches("Topic")), .funs = funs(if_else(is.na(.), 0, .))) %>% # give a value of 0 to the topics of the courses missing from the topic model.
  mutate_at(vars(matches("Topic")), .funs = funs(Grade/10 * .)) %>% # weigh by grade

  group_by(`Student ID`, time) %>%
  summarize_at(vars(matches("Topic")), sum) %>%

  group_by(`Student ID`) %>%
  arrange(time) %>%

  mutate_at(vars(matches("Topic")), lag, default = 0) %>%
  mutate_at(vars(matches("Topic")), cumsum) %>% # topic profile at beginning of period

  ungroup %>%

  arrange(`Student ID`)
```

Raphael's topic profile is coherent: topics with a high value correspond to my academic focus and topics with a low score to themes I never covered (i.e. law, literature).
#Save
```{r save student_TP}
save(student_TP, file = "./App/Recommender System/student_TP.RDATA")
```

# Student GPA

```{r grade_ECTS}
d_transcript_augmented <- d_transcript_augmented %>%

  mutate(is_HUM = `Course ID` %>% str_detect("HUM|SAH"),
         is_SCI = `Course ID` %>% str_detect("SCI|SAS"),
         is_SSC = `Course ID` %>% str_detect("SSC|SAC"),
         is_COR = `Course ID` %>% str_detect("COR"),
         is_SKI = `Course ID` %>% str_detect("SKI|LAN"),
         is_PRO = `Course ID` %>% str_detect("PRO|UGR|CAP")) %>%

  mutate(grade_HUM = if_else(is_HUM, Grade, NA_real_),
         grade_SCI = if_else(is_SCI, Grade, NA_real_),
         grade_SSC = if_else(is_SSC, Grade, NA_real_),
         grade_COR = if_else(is_COR, Grade, NA_real_),
         grade_SKI = if_else(is_SKI, Grade, NA_real_),
         grade_PRO = if_else(is_PRO, Grade, NA_real_)) %>%

  mutate(ECTS = case_when(
    `Course ID` %>% str_detect("HUM|SCI|SSC|COR|PRO|SA|EXT") ~ 5  ,
    `Course ID` %>% str_detect("SKI|LAN"                   ) ~ 2.5,
    `Course ID` %>% str_detect("CAP|UGR"                   ) ~ 10 ,
    TRUE                                                     ~ 5  )
    ) %>%

  mutate(ECTS_HUM = if_else(is_HUM, ECTS, NA_real_),
         ECTS_SCI = if_else(is_SCI, ECTS, NA_real_),
         ECTS_SSC = if_else(is_SSC, ECTS, NA_real_),
         ECTS_COR = if_else(is_COR, ECTS, NA_real_),
         ECTS_SKI = if_else(is_SKI, ECTS, NA_real_),
         ECTS_PRO = if_else(is_PRO, ECTS, NA_real_)) %>%

  mutate(grade_ECTS     = Grade     * ECTS,
         grade_ECTS_HUM = grade_HUM * ECTS_HUM,
         grade_ECTS_SCI = grade_SCI * ECTS_SCI,
         grade_ECTS_SSC = grade_SSC * ECTS_SSC,
         grade_ECTS_COR = grade_COR * ECTS_COR,
         grade_ECTS_SKI = grade_SKI * ECTS_SKI,
         grade_ECTS_PRO = grade_PRO * ECTS_PRO)
```

```{r student_GPA, cache = TRUE}
student_GPA <- d_transcript_augmented %>%

  group_by(`Student ID`, time) %>%
  summarize_at(vars(matches("ECTS")), list) %>%

  group_by(`Student ID`) %>%
  arrange(time) %>%

  mutate_at(vars(matches("ECTS")), purrr::accumulate, c) %>%

  mutate_at(vars(matches("ECTS")), map_dbl, sum, na.rm = TRUE) %>%

  mutate(GPA     = grade_ECTS     / ECTS,
         GPA_HUM = grade_ECTS_HUM / ECTS_HUM,
         GPA_SCI = grade_ECTS_SCI / ECTS_SCI,
         GPA_SSC = grade_ECTS_SSC / ECTS_SSC,
         GPA_COR = grade_ECTS_COR / ECTS_COR,
         GPA_SKI = grade_ECTS_SKI / ECTS_SKI,
         GPA_PRO = grade_ECTS_PRO / ECTS_PRO) %>%

  mutate_at(vars(matches("GPA")), lag) %>%

  mutate_at(vars(matches("GPA")), ~ if_else(is.na(.), mean(d_transcript_augmented$Grade), .)) %>% # substitute missing GPA with the mean GPA across all courses (GPA_HUM is missing if student has no taken any HUM course yet)

  arrange(`Student ID`) 
```

# Join student TP and GPA
```{r student_profile, cache = TRUE}
student_profile <- d_transcript_augmented %>%

  left_join(student_GPA, by = c("Student ID", "time")) %>%

  left_join(student_TP , by = c("Student ID", "time"))
```

```{r}
student_profile_nest <- student_profile %>%

  nest(.key = profile, matches("GPA|Topic")) %>%

  mutate(profile = profile %>% map(as.matrix))
```

# Regressing grade on TP and GPA

```{r find_df}
find_df <- function(course)  student_profile %>% filter(`Course ID` == course)
```

## Lasso
```{r my_cv.glmnet}
my_cv.glmnet <- function(df, alpha = 1, predictors){

  df <- df %>% select(Grade, matches(predictors))

  y <- df[names(df) == "Grade"] %>% as.matrix
  x <- df[names(df) != "Grade"] %>% as.matrix

  nfold <- min(10, floor(nrow(df) / 3))

  cv.glmnet(x, y, nfolds = nfold, type.measure = "mae", alpha = alpha)

}
```

```{r fit_lasso, cache = TRUE}
fit_lasso <- tibble(target = course_target) %>%

  mutate(d = target %>% map(find_df),
         n = d      %>% map_dbl(nrow)) %>%

  filter(n > 20) %>%

  mutate(cv = d %>% map(my_cv.glmnet, predictors = "GPA|Topic")) %>%

  select(-d)
```

```{r fit_lasso extract}
fit_lasso <- fit_lasso %>%

  # Results from CV
  mutate(m_lasso      = cv %>% map    (~ .[["glmnet.fit"]]),
         lambda_min   = cv %>% map_dbl(~ .[["lambda.min"]]),
         lambda_1se   = cv %>% map_dbl(~ .[["lambda.1se"]]),
         index_best   = cv %>% map_dbl(~ which.min(.[["cvm"]])),
         cv_error     = list(cv, index_best) %>% pmap_dbl(~ ..1[["cvm"]][..2]),
         cv_error_sd  = list(cv, index_best) %>% pmap_dbl(~ ..1[["cvsd"]][..2])
         ) %>%

  # Best model
  mutate(

    intercept  = list(m_lasso, index_best) %>% pmap_dbl(~ ..1[["a0"]][..2]),
    coefi      = list(m_lasso, lambda_min) %>% pmap(coef), # clearer output than pmap_dbl(~ ..1[["beta"]][,..2])
    df         = list(m_lasso, index_best) %>% pmap_dbl(~ ..1[["df"]][..2]),

    dev_null   = m_lasso %>% map_dbl(~ .[["nulldev"]]),
    dev_left   = list(m_lasso, index_best) %>% pmap_dbl(~deviance.glmnet(..1)[..2]),
    dev_unexpl = dev_null - dev_left,
    dev_ratio  = list(m_lasso, index_best) %>% pmap_dbl(~ ..1[["dev.ratio"]][..2])

  ) %>%

  select(- m_lasso) %>%

  arrange(cv_error)
```

```{r}
save(fit_lasso, file = "Output/CV.RDATA")
```

```{r}
hist(fit_lasso$cv_error)
mean(fit_lasso$cv_error); weighted.mean(fit_lasso$cv_error, fit_lasso$n)
```

Topic chosen by model are related to the course.
```{r}
beta <- app_model$Beta %>% .[[1]] %>% group_by(topic) %>% top_n(15, beta)
#beta <- topic_model_gb$b_overview[[15]] %>% group_by(topic) %>% top_n(15, beta)

i <- 2; fit_lasso$coefi[[i]]; fit_lasso$target[[i]] # Contemporary Classical Sociology: topic ~ law, foreign policy, culture
beta %>% filter(topic == "Topic 18")
beta %>% filter(topic == "Topic 8" )

i <- 12; fit_lasso$coefi[[i]]; fit_lasso$target[[i]] # Business Administration: topic ~ business
beta %>% filter(topic == "Topic 19")

i <- 3; fit_lasso$coefi[[i]]; fit_lasso$target[[i]] # Culture, Politics and SOciety in Contemporary Asia: topic ~ law, international, policy
beta %>% filter(topic == "Topic 21") 
beta %>% filter(topic == "Topic 23")

i <- 9; fit_lasso$coefi[[i]]; fit_lasso$target[[i]] # history of Western Political Thought: GPA_HUM and topic ~ natural science
beta %>% filter(topic == "Topic 7") 
```

# Preparatory courses

```{r}
fit_lasso_topic <- tibble(target = course_target) %>%

  mutate(d = target %>% map(find_df),
         n = d      %>% map_dbl(nrow)) %>%

  filter(n > 20) %>%

  mutate(cv = d %>% map(my_cv.glmnet, alpha = 0, predictors = "Topic")) %>%

  select(-d)
```

```{r}
fit_lasso_topic <- fit_lasso_topic %>%

  # Results from CV
  mutate(m_lasso      = cv %>% map    (~ .[["glmnet.fit"]]),
         lambda_min   = cv %>% map_dbl(~ .[["lambda.min"]]),
         index_best   = cv %>% map_dbl(~ which.min(.[["cvm"]]))
         ) %>%

  select(- cv) %>%

  # Best model
  mutate(coefi_v = list(m_lasso, index_best) %>% pmap(~ ..1[["beta"]][,..2]),
         coefi_v = coefi_v %>% map(~ tibble(topic = names(.), weight = .)),
         coefi_m = list(m_lasso, lambda_min) %>% pmap(coef)) %>%

  # extract coefficients
  unnest(coefi_v) %>%

  select(target, topic, weight)

print(fit_lasso_topic)
```

```{r}
d_prep <- fit_lasso_topic %>%

  left_join(gamma, by = "topic") %>%
  rename(`Preparatory Courses` = document) %>%
  filter(target != `Preparatory Courses`) %>%

  # exclude 3000 (advanced) courses from preparation courses
  filter(! `Preparatory Courses` %>% str_detect("^...30")) %>%

  # double prep_score of intro courses
  mutate(is_intro = `Preparatory Courses` %>% str_detect("^...10")) %>%
  mutate(weight = weight * (1 + is_intro)) %>%

  # prep_score
  mutate(prep_score = weight * gamma) %>%

  # aggregate prep_score
  group_by(target, `Preparatory Courses`) %>%
  summarize(prep_score = sum(prep_score)) %>%

  # identify top 15 preparation courses for each course
  group_by(target) %>%
  top_n(15, prep_score) %>%
  arrange(target, desc(prep_score))

print(d_prep)
```

# Save

```{r}
fit_lasso_app <- fit_lasso %>% select(target, cv)

student_profile_nest_app <- student_profile_nest %>%

  group_by(`Student ID`) %>% # only keep most recent profile of each student
  arrange(desc(time)) %>%
  slice(1) %>%

  select(`Student ID`, profile)

save(fit_lasso_app, student_profile_nest_app, d_prep, file = "APP/Recommender System/grade_prediction.RDATA")
```

# Predict

```{r my_predict}
my_predict <- function(model, profile){

  predict.cv.glmnet(object = model, newx = profile, s = "lambda.min")

}
```

```{r}
course_ID  <- c("COR1005", "HUM2005") # input$course
student_ID <- "6087587" # input$student

student_prof <- student_profile_nest_app %>%

  filter(`Student ID` == student_ID) %>%

  pull(profile) %>% .[[1]]

fit_lasso_app %>%

  filter(target %in% course_ID) %>%

  mutate(prediction = cv %>% map_dbl(my_predict, student_prof)) %>%

  mutate(flag_red    = prediction < 5.5,
         flag_orange = prediction %>% between(5.5, 7),
         flag_green  = prediction > 7)
```

# Extra: find best alpha

Best alpha turns out to be 1 (lasso). Good, because predictor regularization also serves as predictor selection.

```{r, cache = T}
n_alpha <- 101
mae     <- numeric(n_alpha)
alphas  <- seq(0, 1, length.out = n_alpha)

for(i in 1 : n_alpha){

  fit_lasso <- tibble(target = course_target) %>%

    mutate(d = target %>% map(find_df),
           n = d      %>% map_dbl(nrow)) %>%

    filter(n > 20) %>%

    mutate(cv = d %>% map(my_cv.glmnet, alpha = alphas[i], predictors = "GPA|Topic")) %>%

    mutate(index_best = cv                   %>% map_dbl (~ which.min(.[["cvm"]])),
           cv_error   = list(cv, index_best) %>% pmap_dbl(~ ..1[["cvm"]][..2]    ))

  mae[i] <- weighted.mean(fit_lasso$cv_error, fit_lasso$n)

}

# slight decrease in mae as alpha increase. Set alpha = 1 in my_cv.glmnet().
m <- lm(mae ~ alphas)
plot(alphas, mae)
abline(m, col = "red")
m %>% summary
```