### 2019-05-15 Pillar 1 - Topic Profile.Rmd
use selected topic model

---
title: "Pillar 1 - Student Topic Profile"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

__Considerations__:

* extract course descriptions from courses not offer in 2018-2019 e.g. SCI2012.

* give more weight to 3000-level courses

* consider following predictors:
     * GPA (per concentration)

# Set up

```{r setup, include = FALSE} 
knitr::opts_chunk$set(cache.path = "Cache/Pillar 1/")

library(magrittr)
library(glmnet)
library(leaps)
library(tidyverse)
```

```{r load, cache.lazy = TRUE} 
load("Output/app_model.RDATA")
load("Output/data_student.RDATA")
load("Output/course_current.RDATA")

course_target <- course_current[! str_detect(course_current, "...10..")]
course_target <- union(course_target, str_subset(course_current, "COR"))# include core courses
```

# Topic Model
```{r gamma_spread}
gamma <- app_model$Gamma[[1]] %>%  mutate(topic = topic %>% str_replace(" ", "_"))
gamma_spread <- gamma %>% spread(topic, gamma)
beta  <- app_model$Beta %>% .[[1]] %>% group_by(topic) %>% top_n(15, beta)
remove(app_model)
```

# Students' Topic Expertise
```{r student_TP} 
student_TP <- d_transcript_augmented %>%

  left_join(gamma_spread, by = c("Course ID" = "document")) %>%

  mutate_at(vars(matches("Topic")), .funs = funs(if_else(is.na(.), 0, .))) %>% # give a value of 0 to the topics of the courses missing from the topic model.
  mutate_at(vars(matches("Topic")), .funs = funs(Grade/10 * .)) %>% # weigh by grade

  group_by(`Student ID`, time) %>%
  summarize_at(vars(matches("Topic")), sum) %>%

  group_by(`Student ID`) %>%
  arrange(time) %>%

  mutate_at(vars(matches("Topic")), lag, default = 0) %>%
  mutate_at(vars(matches("Topic")), cumsum) %>% # topic profile at beginning of period

  ungroup %>%

  arrange(`Student ID`)
```

Raphael's topic profile is coherent: topics with a high value correspond to my academic focus and topics with a low score to themes I never covered (i.e. law, literature).

# Students' Academic Performance

```{r grade_ECTS} 
d_transcript_augmented <- d_transcript_augmented %>%

  mutate(is_HUM = `Course ID` %>% str_detect("HUM|SAH"),
         is_SCI = `Course ID` %>% str_detect("SCI|SAS"),
         is_SSC = `Course ID` %>% str_detect("SSC|SAC"),
         is_COR = `Course ID` %>% str_detect("COR"),
         is_SKI = `Course ID` %>% str_detect("SKI|LAN"),
         is_PRO = `Course ID` %>% str_detect("PRO|UGR|CAP")) %>%

  mutate(grade_HUM = if_else(is_HUM, Grade, NA_real_),
         grade_SCI = if_else(is_SCI, Grade, NA_real_),
         grade_SSC = if_else(is_SSC, Grade, NA_real_),
         grade_COR = if_else(is_COR, Grade, NA_real_),
         grade_SKI = if_else(is_SKI, Grade, NA_real_),
         grade_PRO = if_else(is_PRO, Grade, NA_real_)) %>%

  mutate(ECTS = case_when(
    `Course ID` %>% str_detect("HUM|SCI|SSC|COR|PRO|SA|EXT") ~ 5  ,
    `Course ID` %>% str_detect("SKI|LAN"                   ) ~ 2.5,
    `Course ID` %>% str_detect("CAP|UGR"                   ) ~ 10 ,
    TRUE                                                     ~ 5  )
    ) %>%

  mutate(ECTS_HUM = if_else(is_HUM, ECTS, NA_real_),
         ECTS_SCI = if_else(is_SCI, ECTS, NA_real_),
         ECTS_SSC = if_else(is_SSC, ECTS, NA_real_),
         ECTS_COR = if_else(is_COR, ECTS, NA_real_),
         ECTS_SKI = if_else(is_SKI, ECTS, NA_real_),
         ECTS_PRO = if_else(is_PRO, ECTS, NA_real_)) %>%

  mutate(grade_ECTS     = Grade     * ECTS,
         grade_ECTS_HUM = grade_HUM * ECTS_HUM,
         grade_ECTS_SCI = grade_SCI * ECTS_SCI,
         grade_ECTS_SSC = grade_SSC * ECTS_SSC,
         grade_ECTS_COR = grade_COR * ECTS_COR,
         grade_ECTS_SKI = grade_SKI * ECTS_SKI,
         grade_ECTS_PRO = grade_PRO * ECTS_PRO)
```

```{r student_GPA, cache = TRUE} 
student_GPA <- d_transcript_augmented %>%

  group_by(`Student ID`, time) %>%
  summarize_at(vars(matches("ECTS")), list) %>%

  group_by(`Student ID`) %>%
  arrange(time) %>%

  mutate_at(vars(matches("ECTS")), purrr::accumulate, c) %>%

  mutate_at(vars(matches("ECTS")), map_dbl, sum, na.rm = TRUE) %>%

  mutate(GPA     = grade_ECTS     / ECTS,
         GPA_HUM = grade_ECTS_HUM / ECTS_HUM,
         GPA_SCI = grade_ECTS_SCI / ECTS_SCI,
         GPA_SSC = grade_ECTS_SSC / ECTS_SSC,
         GPA_COR = grade_ECTS_COR / ECTS_COR,
         GPA_SKI = grade_ECTS_SKI / ECTS_SKI,
         GPA_PRO = grade_ECTS_PRO / ECTS_PRO) %>%

  mutate_at(vars(matches("GPA")), lag) %>%

  mutate_at(vars(matches("GPA")), ~ if_else(is.na(.), mean(d_transcript_augmented$Grade), .)) %>% # substitute missing GPA with the mean GPA across all courses (GPA_HUM is missing if student has no taken any HUM course yet)

  arrange(`Student ID`) 
```

# Join student TP and GPA
```{r student_profile, cache = TRUE} 
student_profile <- d_transcript_augmented %>%

  left_join(student_GPA, by = c("Student ID", "time")) %>%

  left_join(student_TP , by = c("Student ID", "time"))
```

Remove course enrolments from 2007 from the model because we have no information on the previous course enrolment of these students, hence no data on their previous academic performance and skills acquired in previous courses.
```{r}
student_profile <- student_profile %>% filter(Year_numerical != 2007)
```

## Save
```{r save student_TP} 
save(student_profile, file = "Output/student_profile.RDATA")
```
# Regressing grade on TP and GPA

```{r find_df} 
find_df <- function(course)  student_profile %>% filter(`Course ID` == course)
```

```{r my_cv.glmnet} 
my_cv.glmnet <- function(df, alpha = 1, predictors){

  df <- df %>% select(Grade, matches(predictors))

  y <- df[names(df) == "Grade"] %>% as.matrix
  x <- df[names(df) != "Grade"] %>% as.matrix

  nfold <- min(10, floor(nrow(df) / 3)) # try nfold = 10

  cv.glmnet(x, y, nfolds = nfold, type.measure = "mae", alpha = alpha)

}
```

## Model fit
```{r fit_lasso, cache = TRUE} 
fit_lasso <- tibble(target = course_target) %>%

  mutate(d = target %>% map(find_df),
         n = d      %>% map_dbl(nrow)) %>%

  filter(n > 20) %>%

  mutate(cv = d %>% map(my_cv.glmnet, predictors = "GPA|Topic")) %>%

  select(-d)
```

## Extract results
```{r fit_lasso extract} 
fit_lasso <- fit_lasso %>%

  # Results from CV
  mutate(m_lasso      = cv %>% map    (~ .[["glmnet.fit"]]),
         lambda_min   = cv %>% map_dbl(~ .[["lambda.min"]]),
         lambda_1se   = cv %>% map_dbl(~ .[["lambda.1se"]]),
         index_best   = cv %>% map_dbl(~ which.min(.[["cvm"]])),
         cv_error     = list(cv, index_best) %>% pmap_dbl(~ ..1[["cvm"]][..2]),
         cv_error_sd  = list(cv, index_best) %>% pmap_dbl(~ ..1[["cvsd"]][..2])
         ) %>%

  # Best model
  mutate(

    intercept  = list(m_lasso, index_best) %>% pmap_dbl(~ ..1[["a0"]][..2]),
    coefi      = list(m_lasso, lambda_min) %>% pmap(coef), # clearer output than pmap_dbl(~ ..1[["beta"]][,..2])
    df         = list(m_lasso, index_best) %>% pmap_dbl(~ ..1[["df"]][..2]),

    dev_null   = m_lasso %>% map_dbl(~ .[["nulldev"]]),
    dev_left   = list(m_lasso, index_best) %>% pmap_dbl(~deviance.glmnet(..1)[..2]),
    dev_unexpl = dev_null - dev_left,
    dev_ratio  = list(m_lasso, index_best) %>% pmap_dbl(~ ..1[["dev.ratio"]][..2])

  ) %>%

  select(- m_lasso) %>%

  arrange(cv_error)
```

```{r} 
save(fit_lasso, file = "Output/CV.RDATA")
```

```{r} 
hist(fit_lasso$cv_error)
mean(fit_lasso$cv_error); weighted.mean(fit_lasso$cv_error, fit_lasso$n)
```

Topic chosen by model are related to the course.
```{r}
i <- 1; fit_lasso$coefi[[i]]; fit_lasso$target[[i]] # Contemporary Classical Sociology: topic ~ law, foreign policy, culture
beta %>% filter(topic == "Topic 27")
beta %>% filter(topic == "Topic 62")

i <- 3; fit_lasso$coefi[[i]]; fit_lasso$target[[i]] # Culture, Politics and SOciety in Contemporary Asia: topic ~ law, international, policy
beta %>% filter(topic == "Topic 29") 
beta %>% filter(topic == "Topic 64")

i <- 5; fit_lasso$coefi[[i]]; fit_lasso$target[[i]] # history of Western Political Thought: GPA_HUM and topic ~ natural science
beta %>% filter(topic == "Topic 16")
beta %>% filter(topic == "Topic 4")
```

# Preparatory courses

```{r} 
fit_lasso_topic <- tibble(target = course_target) %>%

  mutate(d = target %>% map(find_df),
         n = d      %>% map_dbl(nrow)) %>%

  filter(n > 20) %>%

  mutate(cv = d %>% map(my_cv.glmnet, alpha = 0, predictors = "Topic")) %>%

  select(-d)
```

```{r} 
fit_lasso_topic <- fit_lasso_topic %>%

  # Results from CV
  mutate(m_lasso      = cv %>% map    (~ .[["glmnet.fit"]]),
         lambda_min   = cv %>% map_dbl(~ .[["lambda.min"]]),
         index_best   = cv %>% map_dbl(~ which.min(.[["cvm"]]))
         ) %>%

  select(- cv) %>%

  # Best model
  mutate(coefi_v = list(m_lasso, index_best) %>% pmap(~ ..1[["beta"]][,..2]),
         coefi_v = coefi_v %>% map(~ tibble(topic = names(.), weight = .)),
         coefi_m = list(m_lasso, lambda_min) %>% pmap(coef)) %>%

  # extract coefficients
  unnest(coefi_v) %>%

  select(target, topic, weight)

print(fit_lasso_topic)
```

```{r} 
d_prep <- fit_lasso_topic %>%

  left_join(gamma, by = "topic") %>%
  rename(`Preparatory Courses` = document) %>%
  filter(target != `Preparatory Courses`) %>%

  # exclude 3000 (advanced) courses from preparation courses
  filter(! `Preparatory Courses` %>% str_detect("^...30")) %>%

  # double prep_score of intro courses
  mutate(is_intro = `Preparatory Courses` %>% str_detect("^...10")) %>%
  mutate(weight = weight * (1 + is_intro)) %>%

  # prep_score
  mutate(prep_score = weight * gamma) %>%

  # aggregate prep_score
  group_by(target, `Preparatory Courses`) %>%
  summarize(prep_score = sum(prep_score)) %>%

  # identify top 15 preparation courses for each course
  group_by(target) %>%
  top_n(15, prep_score) %>%
  arrange(target, desc(prep_score))

print(d_prep)
```

# Save

```{r} 
fit_lasso_app <- fit_lasso %>% select(target, cv)

student_profile_nest_app <- student_profile %>%

  nest(.key = profile, matches("GPA|Topic")) %>%

  mutate(profile = profile %>% map(as.matrix)) %>%

  group_by(`Student ID`) %>% # only keep most recent profile of each student
  arrange(desc(time)) %>%
  slice(1) %>%

  select(`Student ID`, profile)

save(fit_lasso_app, student_profile_nest_app, d_prep, file = "APP/Recommender System/grade_prediction.RDATA")
```

# Predict

```{r my_predict} 
my_predict <- function(model, profile){

  predict.cv.glmnet(object = model, newx = profile, s = "lambda.min")

}
```

```{r} 
course_ID  <- c("COR1005", "HUM2005") # input$course
student_ID <- "6087587" # input$student

student_prof <- student_profile_nest_app %>%

  filter(`Student ID` == student_ID) %>%

  pull(profile) %>% .[[1]]

fit_lasso_app %>%

  filter(target %in% course_ID) %>%

  mutate(prediction = cv %>% map_dbl(my_predict, student_prof)) %>%

  mutate(flag_red    = prediction < 5.5,
         flag_orange = prediction %>% between(5.5, 7),
         flag_green  = prediction > 7)
```

# Extra: find best alpha

Best alpha turns out to be 1 (lasso). Good, because predictor regularization also serves as predictor selection.

```{r, cache = T} 
n_alpha <- 101
mae     <- numeric(n_alpha)
alphas  <- seq(0, 1, length.out = n_alpha)

for(i in 1 : n_alpha){

  fit_lasso <- tibble(target = course_target) %>%

    mutate(d = target %>% map(find_df),
           n = d      %>% map_dbl(nrow)) %>%

    filter(n > 20) %>%

    mutate(cv = d %>% map(my_cv.glmnet, alpha = alphas[i], predictors = "GPA|Topic")) %>%

    mutate(index_best = cv                   %>% map_dbl (~ which.min(.[["cvm"]])),
           cv_error   = list(cv, index_best) %>% pmap_dbl(~ ..1[["cvm"]][..2]    ))

  mae[i] <- weighted.mean(fit_lasso$cv_error, fit_lasso$n)

}

# slight decrease in mae as alpha increase. Set alpha = 1 in my_cv.glmnet().
m <- lm(mae ~ alphas)
plot(alphas, mae)
abline(m, col = "red")
m %>% summary
```