### 2019-05-16 Pillar 1 - Conformal Prediction.Rmd

---
title: "Pillar 1 - Conformal Prediction"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  pdf_document:
    number_sections: true
editor_options: 
  chunk_output_type: console
---

# Set up

```{r setup, include = FALSE}
set.seed(1)

knitr::opts_chunk$set(fig.width = 3, fig.height = 2,
                      echo = FALSE, warning = FALSE)

library(tidyverse)
library(glmnet)

load("Output/CV.RDATA")
load("Output/data_student.RDATA")
load("Output/student_profile.RDATA")
```

# Course selection
I select courses with large and small n, and large and small CV error. I do not select courses whose lasso model contains less than 3 non-zero coefficients (df).

* COR1002: large n
* COR1004: large n and smaller CV error (mean absolute error) than COR1002

* SCI3003: small n and large CV error
* SCI2040: small n and small CV erro

* SSC3044: small CV error
* SSC3038: small CV error and n twice as large as SSC3044

* SCI2018: large CV error
* SCI2010: large CV error and n twice as large as SCI2018

```{r}
course_selection <- c("COR1002", "COR1004", "SCI3003", "SCI2040", "SSC3044", "SSC3038", "SCI2018", "SCI2010")

fit_lasso %>% 
  filter(target %in% course_selection) %>% 
  select(target, n, cv_error, df)
```

# Conformal Prediction setp by step

## Cross-validation

```{r CV helper}
create_Z <- function(course){

  student_profile %>%
    filter(`Course ID` == course) %>%
    select(Y = Grade, matches("GPA|Topic"))

}

add_fold <- function(df, n_fold = 10){

  n         <- nrow(df)
  fold_size <- ceiling(n / n_fold)
  folds     <- rep(1 : 10, fold_size) %>% sample %>% .[1 : n]

  df %>% mutate(fold = folds)

}
```

## Training, callibration and test sets

```{r split helper}
create_test <- function(df, fold_current) df %>% filter(fold == fold_current)

create_training_temp <- function(df, fold_current) df %>% filter(fold != fold_current)

add_allocation <- function(df){

  n <- nrow(df)
  n_training      <- floor(0.67 * n) # 2/3
  n_callibration  <- n - n_training  # 1/3

  allocation <- rep(x     = c("training", "callibration"), 
                    times = c(n_training, n_callibration)) %>%
  sample

  df %>% mutate(allocation = allocation)

}

create_training     <- function(df) df %>% filter(allocation == "training")

create_callibration <- function(df) df %>% filter(allocation == "callibration")
```

## Fit lasso on training set
```{r,lasso helper}
fit_target_model <- function(df){

  X <- df %>% select(matches("GPA|Topic"))
  Y <- df %>% pull(Y)

  cv.glmnet(x = as.matrix(X), y = Y, type.measure = "mae") # lasso by default

}

predict_grade <- function(df, model){

  X <- df %>% select(matches("GPA|Topic"))

  Y_hat <- predict.cv.glmnet(model, newx = as.matrix(X), s = "lambda.min")

  df %>% mutate(Y_hat = as.vector(Y_hat),
                error = abs(Y_hat - Y))

}

fit_error_model <- function(df){

  X <- df %>% select(matches("GPA|Topic"))
  error <- df %>% pull(error)

  cv.glmnet(x = as.matrix(X), y = error, type.measure = "mae") # lasso by default

}

predict_error <- function(df, model){

  X <- df %>% select(matches("GPA|Topic"))

  error_hat <- predict.cv.glmnet(model, newx = as.matrix(X), s = "lambda.min")

  df %>% mutate(error_hat = as.vector(error_hat))

}
```

## Compute non-conformity scores on callibration test
```{r}
compute_ncs <- function(df) df %>% mutate(alpha = abs(Y - Y_hat) / error_hat )

compute_confidence <- function(df) df %>% mutate(confidence = rank(alpha) / nrow(.) )
```

## Evaluate coverage on test set

```{r}
compute_width <- function(df, confidence_given){

  df <- df %>% 
    distinct(alpha, confidence) %>%
    filter(confidence >= confidence_given)

  min(df$alpha)

}

compute_interval <- function(df, ncs){

  df %>% mutate(
    width       = ncs * error_hat,
    border_low  = Y_hat - width,
    border_high = Y_hat + width
    )
}

compute_hit <- function(df){
  df %>% mutate(hit = (border_low <= Y) & (Y <= border_high))
}
```

# Loop through all courses

```{r, echo = FALSE}
for(course in course_selection){

  # Set up
  callibration <- list()
  coverage     <- list()

  confidence <- seq(0.01, 0.99, 0.01)
  df_coverage <- tibble::tibble(
    confidence = confidence,
    width_mean = NA,
    coverage   = NA
    )

  # data
  Z <- create_Z(course)
  Z <- add_fold(Z)

  for(iteration in 1 : 10){

    # training, callibration and test sets
    Z_test          <- create_test(Z, fold_current = iteration)
    Z_training_temp <- create_training_temp(Z, fold_current = iteration) %>% add_allocation
    Z_training      <- create_training(Z_training_temp)
    Z_callibration  <- create_callibration(Z_training_temp)

    # fit target and error model to training set
    model_target <- fit_target_model(Z_training)
    Z_training   <- predict_grade(Z_training, model_target)
    model_error  <- fit_error_model(Z_training)

    # non-conformity scores on callibration
    Z_callibration <- Z_callibration %>% 
      predict_grade(model_target) %>%
      predict_error(model_error) %>%
      compute_ncs %>%
      compute_confidence

    # coverage on test set
    Z_test <- Z_test %>% 
      predict_grade(model_target) %>%
      predict_error(model_error)

    for(i in 1 : nrow(df_coverage)){ # loop through confidence levels

      confidence_i <- df_coverage$confidence[i]
      ncs          <- compute_width(Z_callibration, confidence_i)

      Z_test <- Z_test %>% compute_interval(ncs = ncs) %>% compute_hit

      coverage_i   <- mean(Z_test$hit)
      width_mean_i <- mean(Z_test$width)

      df_coverage$width_mean[i] <- width_mean_i
      df_coverage$coverage[i]   <- coverage_i

    }

    # Output fold
    callibration[[iteration]] <- Z_callibration %>% mutate(iteration = iteration) %>% select(iteration, confidence, alpha)
    coverage[[iteration]]     <- df_coverage %>% mutate(iteration = iteration)

  } # end CV fold

  # Output

  callibration <- bind_rows(callibration) %>% 
    mutate(iteration = as.factor(iteration))
  coverage     <- bind_rows(coverage) %>%
    mutate(iteration = as.factor(iteration))

  title_course <-  str_c(course, " (n = ", nrow(Z), ")")
  subtitle     <- "A line per CV test sets"

  print(ggplot(data = coverage) +
          geom_hline(yintercept = 1, linetype = "dashed") +
          geom_line(aes(confidence, width_mean, group = iteration, col = iteration), size = .5) +
          labs(title = title_course,
               subtitle = subtitle) +
    theme(legend.position = "none"))

  print(ggplot(data = coverage) +
          geom_abline(slope = 1, linetype = "dashed") +
          geom_line(aes(confidence, coverage, group = iteration, col = iteration), size = .5) +
          labs(title = title_course,
               subtitle = subtitle) +
    theme(legend.position = "none"))
}
```

Illustration of the effect of error_hat on interval's width (for confidence of 99%):
```{r, fig.height=4}
Z_test %>% select(Y, Y_hat, error, error_hat, width, border_low, border_high, hit)
plot(Z_test$error, Z_test$error_hat)
```

Correlation between the actual errors and the predicted errors (error_hat):
```{r}
cor(Z_test$error, Z_test$error_hat)
```

```{r, echo = FALSE, eval = F}
ggplot(data = Z_callibration) + geom_histogram(aes(Y), bins = 15)

ggplot(data = Z_callibration) + geom_histogram(aes(Y_hat), bins = 15)

ggplot(data = Z_callibration) + geom_histogram(aes(alpha), bins = 15)
```

```{r, echo = FALSE, eval = F}
ggplot(data = Z_callibration) + geom_point(aes(Y, Y_hat))

ggplot(data = Z_callibration) + geom_point(aes(Y, alpha))

ggplot(data = Z_callibration) + geom_point(aes(Y_hat, alpha))
```