### 2019-06-05 Useful Functions.Rmd
create script of useful functions to reduce copy-pasting

---
title: "Pillar 1 - Student Topic Profile"
author: "DARS"
date: "`r Sys.Date()`"
output: 
  github_document:
    toc: TRUE
editor_options: 
  chunk_output_type: console
---

# Helper Functions

```{r}
collapse_or       <- function(string) string %>% str_c(collapse = "|")
collapse_space    <- function(string) string %>% str_c(collapse = " ")
```

```{r my_cv.glmnet} 
my_cv.glmnet <- function(df, alpha = 1, predictors){

  df <- df %>% select(Grade, matches(predictors))

  y <- df %>% select(Grade             ) %>% as.matrix
  x <- df %>% select(matches(predictors)) %>% as.matrix

  nfold <- min(10, floor(nrow(df) / 3))

  cv.glmnet(x, y, nfolds = nfold, type.measure = "mae", alpha = alpha)

}
```

# Save

```{r}
save(list = ls(), file = "Output/useful functions.RDATA")
```