### 2019-06-07 Data Preparation - Course Data.Rmd
Raphaels files from pull

version https://git-lfs.github.com/spec/v1
oid sha256:3c135a6e599db989adc786cae3f0b15798f02d897be1bb2c1ac7f8b3f3ce2be0
size 13008