### 2019-06-07 Conference/COPA2019/paper/draft.tex
conferences files

version https://git-lfs.github.com/spec/v1
oid sha256:2433448a17d2e226d09ca4e30b2aff49952f1cde0390349985ba7fb08dceb964
size 38682