### 2019-06-27 Conference/COPA2019/paper/draft.tex

version https://git-lfs.github.com/spec/v1
oid sha256:b3b51fd04d7cb0476a10ca059f2513cef246efc46e81eeb79b8188f333f8975b
size 38884