### 2019-06-27 Conference/COPA2019/paper/draft.tex

version https://git-lfs.github.com/spec/v1
oid sha256:0a71b647a5cbec27061e1261ce040d3449937001d6c29b653305e4d754e6444b
size 287